### 2015-05-18 Harris_2015/create_data.Rmd
Add Harris_2015 analyses

# Setup

```{r, cache = TRUE}
devtools::load_all()
species_range = c(5, 10, 20)        # Numbers of species to evaluate
site_range = c(20, 100, 500, 2500)  # Numbers of sites to evaluate
reps = 24                           # Number of times to repeat combination of sites and species
min_gibbs = 1000                    # Always perform at least this many Gibbs samples

n_env = 2                           # Number of environmental variables in the model
```

# Functions

A function for simulating known $\alpha$ and $\beta$ coefficients.

```{r}
make_coefficients = function(n_spp){

  # Most of the n-choose-2 interactions are assumed to be weak; a few
  # are much stronger.
  true_beta_magnitudes = rexp(choose(n_spp, 2))

  # On average, three-quarters of the simulated interactions are
  # negative.
  b = true_beta_magnitudes * sample(
    c(-1, 1), 
    size = length(true_beta_magnitudes), 
    prob = c(.75, .25),
    replace = TRUE
  )

  # Alpha coefficients are normally distributed
  a = rnorm(n_spp, -1)

  # rosalia stores betas in the upper triangle of a matrix and alphas
  # along the diagonal.
  c(upper = b, diag = a)
}
```

# Simulate and save the simulated data

```{r, cache = TRUE}
for(n_spp in species_range){
  cat("\n", n_spp, "\n")

  for(n_sites in site_range){
    cat("\n")
    for(rep in 1:reps){
      cat(".")

      # Make coefficients using the function above
      par = make_coefficients(n_spp)

      # Save the "true" betas
      truth = par[1:choose(n_spp, 2)]

      # The "true" alphas will only be used during simulation
      alpha = par[-(1:choose(n_spp, 2))]

      # Start beta empty, fill in upper triangle, then fill in lower tri with its transpose
      # to make a symmetric matrix of beta coefficients
      beta = matrix(0, n_spp, n_spp)
      beta[upper.tri(beta)] = truth
      beta = beta + t(beta)

      # Environmental states, simulated from a Gaussian for each site and variable
      env = matrix(rnorm(n_sites * n_env), ncol = n_env)

      # Species only respond to the environment in the second half of the simulations
      sd = 2 * (rep > reps/2)
      alpha_env = matrix(rnorm(n_spp * n_env, sd = sd), nrow = n_env)

      # Simulate the landscape from known process
      i = 0 # Gibbs iteration counter starts at zero
      give_up = FALSE # don't give up simulating unless needed
      incomplete = TRUE # Sampling isn't complete yet

      # Landscape starts as if betas were all zero
      x = matrix(
        plogis(rep(1, n_sites) %*% t(alpha) + env %*% alpha_env), 
        nrow = n_sites, 
        ncol = n_spp
      )

      # Gibbs sampling until enough iterations have passed *and* no species are totally absent
      while(incomplete & !give_up){
        i = i + 1
        for(j in 1:n_spp){
          x[,j] = rbinom(
            nrow(x),
            size = 1,
            prob = plogis(x %*% beta[ , j] + alpha[j] + env %*% alpha_env[,j])
          )
        }

        # Simulation is "incomplete"" if insufficient Monte Carlo samples have been collected
        # or if one or more species are missing from the landscape.
        incomplete = i < min_gibbs | min(colSums(x)) == 0

        # If, after 100*minimum iterations, no valid landscapes have been produced, give
        # up on this set of coefficients.
        give_up = i > min_gibbs * 100 
      }

      if(give_up){
        print("giving up")
        next
      }else{
        file_stem = paste(n_spp, n_sites, rep, sep = "-")

        # Save the simulated landscape in the fakedata folder
        write.csv(
          x, 
          file = paste0("fakedata/matrices/", file_stem, ".csv")
        )

        # Ulrich's "Pairs" software rejects empty ("degenerate") sites
        x_subset = x[rowSums(x) != 0, colSums(x) != 0]

        # "Pairs" expects the data to be transposed relative to the other three
        # methods' inputs.
        write.table(
          t(x_subset), 
          file = paste0("fakedata/matrices/", file_stem, "-transposed.txt"), 
          quote = FALSE
        )

        # save the "true" beta values
        write(
          truth, 
          file = paste0("fakedata/truths/", file_stem, ".txt"), 
          ncolumns = length(truth)
        )
      }
    }
  }
}
```

# Housekeeping

In order to run the Pairs analyses more quickly, I moved the simulated data matrices
into 8 folders to run simultaneously on 8 processor cores.

```{r}
transposed_files = dir("fakedata/matrices", pattern = "transposed", full.names = FALSE)

rep_num = as.integer(gsub(".*-.*-(.*)-transposed.txt", "\\1", transposed_files))

core_num = (rep_num %%8) + 1
for(i in 1:8){
  dir.create(paste0("fakedata/matrices/", i), showWarnings = FALSE)
}

file.rename(
  paste0("fakedata/matrices/", transposed_files), 
  paste0("fakedata/matrices/", core_num, "/", transposed_files)
)

for(i in 1:8){
  write(
    paste0(
      c("comment", sample(dir(paste0("fakedata/matrices/", i), pattern = "transposed"))), 
      collapse = "\r\n"
    ), 
    file = paste0("fakedata/matrices/", i, "/", "list.txt")
  )
}

# Include the .exe file in each folder so it can run on the corresponding files
for(i in 1:8){
  file.copy("fakedata/matrices/pairs.exe", paste0("fakedata/matrices/", i, "/", "pairs.exe"))
}
```