### 2015-05-18 Harris_2015/models.Rmd

---
output: html_document
---
# Simulating landscapes from known coefficients

The following function draws random coefficients for a Markov network of a pre-specified size.
The probability distribution was chosen so that most species pairs interact weakly, a few
act strongly, and that most of the interactions were negative rather than positive.

```{r, cache=TRUE}
make_coefficients = function(n_spp){
  # Exponential distribution has lots of mass near 0 but has
  # a long tail.
  true_beta_magnitudes = rexp(choose(n_spp, 2))

  # Species interactions are negative 3/4 of the time in 
  # these simulations.
  b = true_beta_magnitudes * sample(
    c(-1, 1), 
    size = length(true_beta_magnitudes), 
    prob = c(.75, .25),
    replace = TRUE
  )

  # Species' intercepts are normally distributed
  a = rnorm(n_spp, -1)

  # Return the simulated values.
  # The rosalia function stores pairwise parameters in the upper
  # triangle of an n-by-n matrix and stores the species intercepts
  # along the diagonal, so these values are named accordingly.
  c(upper = b, diag = a)
}
```

I used the following function to generate "true" parameters with the method above, 
simulate a presence-absence landscape based on those parameters using Markov chain Monte Carlo,
and save the results to a "fakedata" folder.

```{r, cache = TRUE}
library(rosalia)

simulate_data = function(n_spp, n_sites, n_rep, min_gibbs = 1000, reps = 24, n_env = 2){

  # Determine the "true" parameters for the simulated assemblage
  par = make_coefficients(n_spp)

  # "True" interaction strengths, to save for later
  truth = par[1:choose(n_spp, 2)]

  # "True" intercepts, possibly adjusted below by environment
  alpha = par[-(1:choose(n_spp, 2))]

  # Turn the interaction values into an n-by-n matrix
  # Start with empty matrix; fill in upper triangle; 
  # then fill in lower triangle with its transpose
  beta = matrix(0, n_spp, n_spp)
  beta[upper.tri(beta)] = truth
  beta = beta + t(beta)

  # Environmental states are normally distributed
  env = matrix(rnorm(n_sites * n_env), ncol = n_env)

  # Species only respond to the environment in the second half of the simulations
  # (i.e. their responses are sampled from a Gaussian with a mean and standard deviation 
  # of zero in the first half)
  sd = 2 * (rep > reps/2)
  alpha_env = matrix(rnorm(n_spp * n_env, sd = sd), nrow = n_env)

  # Simulate the landscape from known process with Gibbs sampling
  i = 0 # Gibbs iteration counter starts at zero
  give_up = FALSE # don't give up simulating unless things run for too long
  incomplete = TRUE # Sampling isn't complete yet

  # Landscape starts as if betas were all zero
  x = matrix(
    plogis(rep(1, n_sites) %*% t(alpha) + env %*% alpha_env), 
    nrow = n_sites, 
    ncol = n_spp
  )

  # Gibbs sampling until enough iterations have passed *and* no species are totally absent
  while(incomplete & !give_up){
    i = i + 1
    for(j in 1:n_spp){
      x[,j] = rbinom(
        nrow(x),
        size = 1,
        prob = plogis(x %*% beta[ , j] + alpha[j] + env %*% alpha_env[,j])
      )
    }

    # If fewer than min_gibbs iterations have passed, or if a species is completely absent,
    # then keep sampling
    incomplete = i < min_gibbs | min(colSums(x)) == 0

    # But give up sampling if the sampling goes on 100x longer than the minimum
    give_up = i > min_gibbs * 100
  }

  if(give_up){
    print("giving up")
    next
  }else{
    # Save the results in a "fake data" folder

    file_stem = paste(n_spp, n_sites, rep, sep = "-")

    # Save the matrix of presence/absence observations
    write.csv(
      x, 
      file = paste0("fakedata/matrices/", file_stem, ".csv")
    )

    # Gotelli and Ulrich's Pairs software rejects empty sites, so I remove them here
    x_subset = x[rowSums(x) != 0, colSums(x) != 0]

    # Gotelli and Ulrich's Pairs method expects the data matrices to be transposed,
    # So I save them separately
    write.table(
      t(x_subset), 
      file = paste0("fakedata/matrices/", file_stem, "-transposed.txt"), 
      quote = FALSE
    )

    # Save the "true" species interactions
    write(
      truth, 
      file = paste0("fakedata/truths/", file_stem, ".txt"), 
      ncolumns = length(truth)
    )
  }
}
```

# Model fitting

## R-compatible methods

Three of the four methods are described in this subsection.  The "Pairs" software
needs to be analyzed outise of R and is described below.

### Rosalia

The `rosalia` function finds (penalized) maximum likelihood estimates for the
$\alpha$ and $\beta$ parameters in a Markov network using the `BFGS` method of
the `optim` function for convex optimization.  For these analyses,
I penalized the likelihood with logistic priors on both sets of parameters to stabilize
the model's estimates. In order to ensure convergence, I increased the number of
`BFGS` iterations from 100 (the default) to 1000.

The following function returns a vector of $\beta$ coefficients.

```{r, cache=TRUE, eval = FALSE}
library("rosalia")

fit_rosalia = function(x){
  fit = rosalia(
    x = x, 
    prior = make_logistic_prior(scale = 1), 
    maxit = 1E3
  )

  # Return the upper triangle of the interaction matrix,
  # which has all the beta values.
  relist(fit$par)$upper
}

```

### Partial covariance

The partial covariance is normally computed as the matrix inverse of
the covariance times `-1`.  The `ridgeS` function adds a regularization parameter,
`lambda`, that stabilizes the estimates and makes it possible to 
calculate a partial covariance matrix even when two species are perfectly 
correlated in the observed data.

```{r, cache=TRUE}
fit_partial = function(x){
  library("rags2ridges") # rags2ridges can be installed from CRAN with 
                         # install.packages("rags2ridges")

  # The effective size of the penalty should shrink as more data becomes available
  # and overwhelms it
  fit = -ridgeS(cov(x), lambda = 1 / (5 * nrow(x)))

  # all pairwise estimtes can be found in the upper triangle of the 
  # partial covariance matrix.
  fit[upper.tri(fit)]
}

```

### Sample covariance

The sample covariance provides a simpler estimate of species' interaction strengths.

```{r, cache=TRUE}
fit_cov = function(x){
  fit = cov(x)
  fit[upper.tri(fit)]
}
```

## Analyzing simulated landscapes with the first three models

This function takes a quoted model name (`"cov"`, `"rosalia"`, or `"partial"`) and applies the
corresponding function above to all the matrices in the `fakedata` folder.

It returns a data frame with information about each species pair's simulated landscape,
the "true" interaction strength between the pair, and a "predicted" value from the 
fitted model.

```{r, cache = TRUE}
fit_model = function(method_name, csv_names){
  fit = get(paste("fit", method_name, sep = "_"))
  out_list = lapply(
    csv_names, 
    function(x){
      file_info = as.integer(strsplit(x, "-|\\.csv")[[1]])

      full_path = paste0("fakedata/matrices/", x)

      truth = read.table(
        paste0(
          "fakedata/truths/",
          paste(file_info, collapse = "-"),
          ".txt"
        )
      )

      data.frame(
        n_spp = file_info[1], 
        n_loc = file_info[2],
        n_rep = file_info[3],
        truth = unlist(truth),
        predicted = fit(read.csv(full_path, row.names = 1)),
        method = method_name,
        stringsAsFactors = FALSE
      )
    }
  )
  do.call(rbind, out_list)
}
```

```{r, cache= TRUE}
cov_results = fit_model("cov", dir("fakedata/matrices"))
```

```{r, cache = TRUE}
partial_results = fit_model("partial", dir("fakedata/matrices"))
```

## Settings for fitting a Pairs model

The Pairs program, written by Dr. Werner Ulrich, can be downloaded from 
[his web site](http://www.keib.umk.pl/pairs/?lang=en) as a Windows executable, but no source
code or scripts are available to include here.

I ran the software with the following options:

* Batch mode
* Sequential swap ("s")
* Printing all pairs ("y")
* C-score co-occurrence measure ("c")
* Default confidence limits (0.05)
* Default iterations (100)
* Maximum of 20 species

## Code for parsing the results from Pairs

The Pairs software produces a text file with two tables of results for each data matrix.

For the analyses in this paper, I needed to match the (negative) Z-scores associated with each
species pair with the "true" interaction parameters used to simulate the corresponding
data matrix.  I used the following code to do so.

The Z-scores, (which are based on pairwise C-scores), have the opposite sign of the other metrics
assessed here, so I multiplied them by negative one after importing them.

```{r, cache=TRUE}

raw = readLines("Pairs-results.txt")

# Pull out the file ID (a sequence of digits separated by dashes) 
# from the lines that begin with ">"
ids = gsub("^.[^0-9]*([0-9]*-[0-9]*-[0-9]*).*$", "\\1", grep("^>", raw, value = TRUE))

splitted_ids = strsplit(ids, "-")

# All the tables in `raw` have "MeanScore" in their headers, but I don't
# want the tables that have "ObsNumber" in them.
header_rows = grep("MeanScore", raw)
wrong_header_rows = grep("ObsNumber", raw)

# Find the start and ending lines for the data tables of interest
starts = header_rows + 1
ends = c(
  # The table of interest ends 3 lines before the next table starts
  wrong_header_rows[-1] - 3,
  length(raw)
)

# Loop through the tables and pull out negative Z-scores for each
# species pair, then line them up with the "truth" values.
# Finally, concatenate the results into a giant data.frame
pairs_results = do.call(
  rbind,
  lapply(
    1:length(ids), 
    function(i){

      # Split the table by whitespace
      split_table = strsplit(raw[starts[[i]]:ends[[i]]], "\\s+")

      # Z-scores are the 14th element of each splitted row
      z_score = sapply(split_table, function(x) as.numeric(x[[14]]))

      # The first species' ID is in the third element of each splitted row
      species_1 = sapply(split_table, function(x) as.numeric(x[[3]]))

      # The second species' ID is in the fourth element of each splitted row
      species_2 = sapply(split_table, function(x) as.numeric(x[[4]]))

      truth = as.numeric(
        strsplit(
          readLines(
            paste0(c("fakedata/truths/", ids[i], ".txt"), collapse = "")
          ),
          " "
        )[[1]]
      )

      # The number of species is the first element of each id
      n_spp = as.numeric(splitted_ids[[i]][1])
      n_loc = splitted_ids[[i]][2] # Number of locations is second
      n_rep = splitted_ids[[i]][3] # Replicate number locations is second

      truthmat = matrix(0, n_spp, n_spp)
      truthmat[upper.tri(truthmat)] = truth
      truthmat = truthmat + t(truthmat)

      # Output a data.frame with the "true" parameters, the negative Z-scores
      # and information about the simulated landscape
      data.frame(
        n_spp = as.integer(n_spp), 
        n_loc = as.integer(n_loc),
        n_rep = as.integer(n_rep),
        truth = truthmat[cbind(species_1, species_2)],
        predicted = -z_score,
        method = "Null",
        stringsAsFactors = FALSE
      )
    }
  )
)

```

As noted in the manuscript, I manually altered one Z-score produced by the Pairs software because
it was such a strong outlier that it completely dominated the regression model described below.

```{r, cache = TRUE}
# The largest predicted Z-score was over 1000; the next largest was about 32
head(sort(pairs_results$predicted, decreasing = TRUE))
pairs_results$predicted[which.max(pairs_results$predicted)] = 32.47
```

# Model evaluation

For each method, I fit a linear regression model through the origin in order to put the different
models' estimates on a common scale.  
In these regressions, the weights associated with each species pair in the regression model are 
inversely proportional to the number of species in a given landscape.  In this way, I ensured that 
each landscape (as opposed to each species pair) contributed equally to the regression.

For each species pair, I recorded the residuals between this regression estimate and the "true"
interaction strength used to simulate the original data.

```{r, cache = TRUE}

results = do.call(
  rbind,
  lapply(
    list(cov_results, partial_results, pairs_results),
    function(x){
      residuals = residuals(
        lm(
          truth ~ predicted + 0, 
          data = x, 
          weights = 1/choose(n_spp, 2),
          y = TRUE
        )
      )
      cbind(x, residuals = residuals)
    }
  )
)

# The second half of the simulated landscapes had an environmental component
results$env = results$n_rep > (max(results$n_rep) / 2)

```

I then calculated $R^2$ for each method on each type of landscape (i.e. landscapes with each level
of richness, each number of sites, and presence/absence of environmental variables)

method as one minus the sum of squared residuals divided by
the squared deviations from zero (e.g. `1 - sum(pairs_fit$residuals^2) / sum(pairs_fit$truth^2)` for
the `pairs_results`).

```{r, cache = TRUE}
library(dplyr)

summarized_results = results %>% group_by_("n_spp", "n_loc", "method", "env") %>% 
  do_(
    ~data.frame(r2 = 1 - sum(.$residuals^2) / sum(.$truth^2))
  )

head(summarized_results) # Show the first 6 rows
```

```{r, cache = FALSE}
library(ggplot2)

basic_plot = ggplot(aes(n_loc, r2, color = method, shape = method), data = summarized_results) + 
  geom_line(size = .5) + 
  geom_point(size = 2, fill = "white") + 
  facet_grid(env~n_spp) + 
  geom_hline(yintercept = 0, size = 1/2) +
  geom_vline(xintercept = 0, size = 1)

scaled_plot = basic_plot +
  scale_shape_manual(values = c(21, 25, 15, 18)) + 
  coord_cartesian(ylim = c(-0.1, 1.01)) + 
  ylab(expression(R^2)) + 
  xlab("Number of sites (log scale)") + 
  scale_x_log10(breaks = c(20, 100, 500, 2500), limits = c(20, 2.5E3)) +
  scale_color_manual(values = c("red3", "purple2", "sienna1", "mediumblue"))

scaled_plot + theme_bw(base_size = 11) + 
  theme(panel.margin = grid::unit(1.25, "lines")) + 
  theme(panel.border = element_blank(), axis.line = element_blank()) + 
  theme(
    panel.grid.minor = element_blank(), 
    panel.grid.major.y = element_line(color = "lightgray", size = 1/4), 
    panel.grid.major.x = element_blank()
  ) + 
  theme(strip.background = element_blank(), legend.key = element_blank()) + 
  theme(plot.margin = grid::unit(c(.01, .01, .75, .1), "lines")) + 
  #theme(legend.text = element_text(size = 11)) + 
  theme(
    axis.title.x = element_text(vjust = -0.2, size = 14), 
    axis.title.y = element_text(angle = 0, hjust = -.1, size = 14)
  )

```