### 2015-05-18 Harris_2015/three_species.Rmd
minor fixes for three_species

# Evaluation of the three species in Figure 1

Load dependencies

```{r}
library(dplyr)
library(magrittr)
library(rosalia)
library(knitr)
library(rosalia)
```

```{r}
set.seed(1)
```

Generate the species interaction matrix from Figure 1.

```{r}
alpha = c(3, 2, 1)
beta = matrix(
  c(0, -3, -3, -3, 0, -1, -3, -1, 0), 
  nrow = 3, 
  dimnames = replicate(2, c("tree", "shrub1", "shrub2"), simplify = FALSE)
)

kable(beta)
```

```{r}
# Simulated community has three species at some of 100 locations
n_spp = 3
n_sites = 100

# Re-organize the coefficients for rosalia
truth = c(beta[upper.tri(beta)], alpha)

# Enumerate the 8 possible species assemblages
# using an internal rosalia function
possibilities = rosalia:::generate_possibilities(n_spp)

# In each of these 8 possible assemblages, which of the 6 possible
# species pairs *both* occur?
possible_cooc = sapply(
  1:2^n_spp,
  function(i){
    tcp = tcrossprod(possibilities[i, ])
    c(tcp[upper.tri(tcp)], diag(tcp))
  }
)
```

Simulate 1000 assemblages of 3 species across 100 sites.

Once a landscape has been simulated, test whether `rosalia`, the sample covariance, and the partial 
covariance got the sign of the shrub-shrub interaction correct.
```{r}
evaluation = sapply(
  1:1000,
  function(i){
    x = rosalia:::simulate_data(
      par = truth, 
      possibilities = possibilities, 
      possible_cooc = possible_cooc, 
      n_sites = n_sites
    )
    if(paste0("fakedata/three/", i, ".txt") %in% dir("fakedata/three", full.names = TRUE)){
      # Do nothing: simulated data was already written to disk
    }else{
      write.table(t(x[rowSums(x) > 0, ]), file = paste0("fakedata/three/", i, ".txt"))
    }

    c(
      cov = cov(x)[2, 3] < 0,
      partial = -solve(cov(x))[2, 3] < 0,
      rosalia = rosalia(x, trace = 0, hessian = FALSE)$par[[3]] < 0
    )
  })

```

After running the "Pairs" software on transposed versions of the files simulated above, 
we can import its results to see how the method performed.

```{r}
Pairs.txt = readLines("fakedata/three/Pairs.txt")

# Find the column that contains the Z-score
z_score_column = grep("Z-Score", Pairs.txt, value = TRUE)[2] %>% 
  strsplit(" +") %>% 
  extract2(1) %>% 
  grep("Z-Score", .)

# Find the right rows of the text file, split them on whitespace, pull out
# the column identified above, and find its sign
pairs_evaluation = Pairs.txt[grep("(Var2.*Var1)|(Var1.*Var2)", Pairs.txt)] %>% 
    strsplit(" +") %>%
    sapply(. %>% extract(z_score_column) %>% as.numeric) %>%
    is_greater_than(0) # Z>0 implies negative interactions because the Z-score
                       # relates to the distribution of C-scores, which describe
                       # the frequency with which species *do not* co-occur
```

Percentage of correct sign estimates:
```{r, eval=FALSE}
mean(pairs_evaluation) * 100

rowMeans(evaluation) * 100
```

```{r, echo = FALSE}
out = c(pairs = mean(pairs_evaluation) * 100, rowMeans(evaluation) * 100)

kable(data_frame(model = names(out), percent_correct = paste(out, "%")))
```