### 2015-05-19 README.Rmd
package stuff

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# rosalia: Exact inference for small binary Markov networks

Santa Rosalia, the patron saint of biodiversity, became a controversial symbol in the 1970's
and 1980's for ecological inferences regarding species' interactions. This package, which 
provides tools from statistical physics for drawing such inferences, is named in her honor.

# Installation:

The easiest way to install `rosalia` is with the `devtools` package, which can be installed from 
the R command line with `install.packages("devtools")`.

Once `devtools` is available, `rosalia` can be downloaded and installed with 
`devtools::install_github("davharris/rosalia")`

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```