### 2015-05-20 Harris_2015/manuscript-materials/Appendices.Rmd
add manuscript materials

# Appendices

## 1. Simlulated community structures

I randomly assigned each species pair an interaction magnitude, $\beta_{ij}$, 
drawn from an exponential distribution with rate parameter 1.  I also randomly
assigned three quarters of the interaction coefficients to be negative, and the
remaining quarter of the coefficients to be positive.

For the simulated landscapes where the abiotic environment was constant across
locations, each species' $\alpha$ coefficient was drawn from a normal
distribution with mean -1 and standard deviation 1.  For the remaining
landscapes, two environmental variables, $x_1$ and $x_2$ were sampled from
independent standard normals with different values for each location. Each
species' response to these two environmental variables was a linear function
of these two environmental variables, with coefficients drawn from normal
distributions with mean 0 and standard deviation 2.  In this way,  species'
$\alpha$ coefficients (and thus their occurrence probabilities) depended on
external environmental factors.

Once the "true" coefficients had been generated for each site on the
landscape, I generated possible landscapes using Gibbs sampling.  In each
round of Gibbs sampling, I cycled through all the species, randomly updating
each one's presence/absence vector in response to its conditional occurrence
probability:

$$p(y_i) = \mathrm{logistic}(\alpha_i + \sum_j\beta_{ij}y_j),$$

where the logistic function is $\frac{1}{1+e^{-x}}$ [@murphy_machine_2012]. After 1000 rounds of
sampling, I continued this procedure until I obtained a landscape matrix where
all of the species occurred at least once. I treated the resulting matrices as
"observed" data for analysis.

## 2. Log-likelihood gradient

The *energy* of a Markov network is defined as

$$E(y;\alpha,\beta) = -\sum_{i}\alpha_i y_i - \sum_{i\neq j}\beta_{ij}y_i y_j.$$

The energy function can be used to define the log-likelihood of a given $y$
vector as

$$\log\mathcal{L}(y;\alpha, \beta) = -E(y;\alpha, \beta) - \log(Z(\alpha, \beta)).$$

Here, $Z(\alpha, \beta)$ is the *partition function*, a scaling factor defined
as

$$Z(\alpha, \beta) = \sum_{y \in Y}e^{-E(y;\alpha, \beta)},$$

where $Y$ is the set of all possible $y$ vectors.

The partial derivative of the log-likelihood with respect to $\alpha_i$ is

$$\frac{\partial}{\partial \alpha_i} \log\mathcal{L}(y;\alpha, \beta) = y_i - p(y_i;\alpha, \beta).$$

The first term, $y_i$, is zero if species $i$ is absent in the observed
assemblage and one if it's present. The latter term, $p(y_i;\alpha, \beta)$,
describes the expected probability of observing species $i$ under the current
values of $\alpha$ and $\beta$. It comes from the derivative of the partition
function, as derived in (learning Boltzmann machines, Murphy 2012, etc.).
Following the gradient of $\alpha_i$ adjusts the expected probability of
observing species $i$ until it matches the observed value and the two terms in
the gradient cancel one another out.

The partial derivative of the log-likelihood with respect to $\beta_{ij}$
can be derived similarly as

$$\frac{\partial}{\partial \alpha_i} \log\mathcal{L}(y;\alpha, \beta) =y_i y_j - p(y_i y_j;\alpha, \beta).$$

Following this gradient adjusts the expected probability of co-occurrence
between species $i$ and species $j$ until this value matches the observed 
co-occurrence frequency and the two terms in the gradient cancel one another out.

# References