### 2015-05-21 Harris_2015/manuscript-materials/Appendices.Rmd
I think this one's actually ready to submit.

## 2. Log-likelihood gradient

The *energy* of a Markov network is defined as

$$E(y;\alpha,\beta) = -\sum_{i}\alpha_i y_i - \sum_{i\neq j}\beta_{ij}y_i y_j.$$

The energy function can be used to define the log-likelihood of a given $y$
vector as

$$\log\mathcal{L}(y;\alpha, \beta) = -E(y;\alpha, \beta) - \log(Z(\alpha, \beta)).$$

Here, $Z(\alpha, \beta)$ is the *partition function*, a scaling factor defined
as

$$Z(\alpha, \beta) = \sum_{y \in Y}e^{-E(y;\alpha, \beta)},$$

where $Y$ is the set of all possible $y$ vectors.

The partial derivative of the log-likelihood with respect to $\alpha_i$ is

$$\frac{\partial}{\partial \alpha_i} \log\mathcal{L}(y;\alpha, \beta) = y_i - p(y_i;\alpha, \beta).$$

The first term, $y_i$, is zero if species $i$ is absent in the observed
assemblage and one if it's present. The latter term, $p(y_i;\alpha, \beta)$,
describes the expected probability of observing species $i$ under the current
values of $\alpha$ and $\beta$. It comes from the derivative of the partition
function, as derived in (learning Boltzmann machines, Murphy 2012, etc.).
Following the gradient of $\alpha_i$ adjusts the expected probability of
observing species $i$ until it matches the observed value and the two terms in
the gradient cancel one another out.

The partial derivative of the log-likelihood with respect to $\beta_{ij}$
can be derived similarly as

$$\frac{\partial}{\partial \alpha_i} \log\mathcal{L}(y;\alpha, \beta) =y_i y_j - p(y_i y_j;\alpha, \beta).$$

Following this gradient adjusts the expected probability of co-occurrence
between species $i$ and species $j$ until this value matches the observed 
co-occurrence frequency and the two terms in the gradient cancel one another out.

# References