### 2015-06-13 stochastic/draft.Rmd
add a script for stochastic approximation

```{r}
library(rosalia)
library(mistnet)
library(progress)
library(beepr)

beeping = TRUE
```

```{r}
`%plus%` = mistnet:::`%plus%`
logistic = binomial()$linkinv
rbern = function(p){rbinom(length(p), size = 1, prob = p)}
```

```{r}
set.seed(1)
n_spp = 250
n_loc = 2500
n_env = 3
n_gibbs = 1000

true_beta_vec = rnorm(
  choose(n_spp, 2), 
  sd = abs(rnorm(choose(n_spp, 2))) / 5
)

true_beta = matrix(0, nrow = n_spp, ncol = n_spp)
true_beta[upper.tri(true_beta)] = true_beta_vec
true_beta = true_beta + t(true_beta)

true_intercepts = rnorm(n_spp)

true_env = matrix(rnorm(n_loc * n_env, 0), nrow = n_loc)

true_alpha_env = matrix(rnorm(n_spp * n_env, 0, 0.5), nrow = n_env)

true_alpha = true_env %*% true_alpha_env
```

```{r, cache = TRUE}
y = matrix(0.5, nrow = n_loc, ncol = n_spp)

pb <- progress_bar$new(
  format = "  simulating landscape [:bar] :percent eta: :eta",
  total = n_gibbs,
  clear = FALSE
)

for(i in 1:n_gibbs){
  pb$tick()
  for(j in sample.int(n_spp)){
    y[,j] = rbern(logistic(true_alpha[ , j] + y %*% true_beta[ , j]))
  }
}

if(beeping){beep()}
```

```{r}
# Calculate sufficient statistics of the data
y_stats = crossprod(y)
y_env_stats = t(true_env) %*% y
```

```{r}
y_sim = matrix(0.5, nrow = nrow(y), ncol = ncol(y))
env = true_env

alpha_env = delta_alpha_env = matrix(0, nrow = n_env, ncol = n_spp)
alpha_species = delta_alpha_species = qlogis(colMeans(y))
beta = delta_beta = matrix(0, nrow = n_spp, ncol = n_spp)
alpha = matrix(0, nrow = n_spp, ncol = n_spp) # no delta alpha to initialize
                                              # b/c alpha not optimized directly
```

```{r}
initial_learning_rate = 0.1
momentum = 0.9

pb <- progress_bar$new(
  format = "  Fitting [:bar] :percent eta: :eta",
  total = n_gibbs,
  clear = FALSE
)

for(i in 1:n_gibbs){
  pb$tick()

  ##############################
  # Gibbs sampling for predicted species composition
  ##############################

  # Update alpha
  alpha = env %*% alpha_env %plus% alpha_species

  # Sample entries in y_sim from their conditional distribution (Gibbs sampling)
  for(j in sample.int(n_spp)){
    y_sim[,j] = rbern(logistic(alpha[ , j] + y_sim %*% beta[ , j]))
  }

  ##############################
  # Stochastic approximation for updating alpha and beta
  ##############################

  # Update learning rate
  learning_rate = initial_learning_rate * 1000 / (999 + i)

  # Calculate sufficient statistics
  y_sim_stats = crossprod(y_sim)
  y_sim_env_stats = t(env) %*% y_sim

  # Calculate the gradient with respect to alpha and beta
  delta_stats = y_stats - y_sim_stats
  beta_grad = (delta_stats - 10 * beta) / 10 / n_loc # Subtract beta as weight decay

  alpha_species_grad = diag(beta_grad) / n_loc
  diag(beta_grad) = 0
  alpha_env_grad = (y_env_stats - y_sim_env_stats - alpha_env)  / n_loc

  # Calculate parameter updates
  delta_beta = beta_grad * learning_rate + momentum * delta_beta
  delta_alpha_species = alpha_species_grad * learning_rate # no momentum
  delta_alpha_env = alpha_env_grad * learning_rate + momentum  * delta_alpha_env

  beta = beta + delta_beta
  alpha_species = alpha_species + delta_alpha_species
  alpha_env = alpha_env + delta_alpha_env

  if(i %% 10 == 0){
    plot(beta[upper.tri(beta)], true_beta[upper.tri(beta)],
         pch = 16, cex = .5, col = "#00000050",
         xlab = "Estimated interaction strength",
         ylab = "\"True\" interaction strength",
         main = paste(
           "R^2 =", 
           signif(cor(beta[upper.tri(beta)], true_beta[upper.tri(beta)])^2, 2)
          )
    )
    abline(v = 0, h = 0)
  }
}
```

```{r}
pdf("poster-beta.pdf", width = 4, height = 4)
plot(beta[upper.tri(beta)], true_beta[upper.tri(beta)],
     pch = 16, cex = .4, col = "#00000020",
     xlab = "Estimated interaction strength",
     ylab = "\"True\" interaction strength",
     main = "Recoverd pairwise interactions",
     las = 1,
     cex.axis = 0.75
    )
text(
  -0.7, 
  1,
  paste(
    "R^2 =", 
    signif(cor(beta[upper.tri(beta)], true_beta[upper.tri(beta)])^2, 2)
  )
)

dev.off()
```

```{r}
pdf("poster-alpha.pdf", width = 4, height = 4)
plot(alpha_env, true_alpha_env,
     pch = 16, cex = .4, col = "#00000090",
     xlab = "Estimated environmental coefficient",
     ylab = "\"True\" environmental coefficient",
     main = "Recovered environmental coefficients",
     las = 1,
     cex.axis = 0.75
)
text(
  -1, 
  1,
  paste(
    "R^2 =", 
    signif(cor(c(alpha_env), c(true_alpha_env))^2, 2)
  )
)
dev.off()

```