### 2015-10-25 Harris_2015/recovery.Rmd
various

---
output: 
  pdf_document:
    keep_tex: true
title: "Appendix 3: Estimating species interactions"
subtitle: \"Estimating species interactions from observational data with Markov networks\"
author: David J. Harris
fontsize: 11pt
---

This document describes how the different models were fit to the simulated data 
from Appendix 2 and how each model's performance was evaluated.[^1]

[^1]: The PDF version of this document has been manually altered to omit 150
lines of output from the `corpcor` package of the form "`## Estimating optimal shrinkage intensity lambda (correlation matrix): 0.0326`"

Initialization:

```{r, cache = FALSE, message = FALSE, eval = TRUE}

mc.cores = 2

library(dplyr)        # For manipulating data structures
library(corpcor)      # For regularized partial covariances
library(rosalia)      # For Markov networks
library(arm)          # For regularized logistic regression
library(BayesComm)    # For joint species distribution modeling
library(parallel)     # for mclapply
set.seed(1)
```

Load in the results from the `pairs` program, run outside of R with
the following options:

  * Batch mode
  * Sequential swap ("s")
  * Printing all pairs ("y")
  * C-score co-occurrence measure ("c")
  * Default confidence limits (0.05)
  * Default iterations (100)
  * Maximum of 20 species

A function to import the data file and run each method on it:

```{r, cache = FALSE, eval = TRUE}
fit_all = function(identifier){
  ######## Import ########

  data_filename = paste0("fakedata/matrices/", identifier, ".csv")
  truth_filename = paste0("fakedata/truths/", identifier, ".txt")

  # first column is row numbers; drop it
  raw_obs = as.matrix(read.csv(data_filename)[ , -1])

  # Identify species that are never present (or never absent) so they 
  # can be dropped
  species_is_variable = diag(var(raw_obs)) > 0
  pair_is_variable = tcrossprod(species_is_variable) > 0

  x = raw_obs[ , species_is_variable]
  truth = unlist(read.table(truth_filename))[pair_is_variable[upper.tri(pair_is_variable)]]

  splitname = strsplit(identifier, "/|-|\\.")[[1]]
  n_spp = as.integer(splitname[[1]]) # simulated species, NOT present species
  n_sites = as.integer(splitname[[2]])
  rep_name = splitname[[3]]

  # Species IDs
  sp1 = combn(colnames(x), 2)[1, ]
  sp2 = combn(colnames(x), 2)[2, ]

  ######## Partial correlations ########
  p_corr = pcor.shrink(x)

  ######## Correlations ########
  corr = cor(x)
  ######## GLM ########
  coef_matrix = matrix(0, ncol(x), ncol(x))
  for(i in 1:ncol(x)){
    if(var(x[,i]) > 0){
      coefs = coef(bayesglm(x[,i] ~ x[ , -i], family = binomial))[-1]
      coef_matrix[i, -i] = coefs
    }
  }
  coef_matrix = (coef_matrix + t(coef_matrix)) / 2

  ######## Markov network ########
  rosie = rosalia(x, maxit = 200, trace = 0, prior = make_logistic_prior(scale = 2), hessian = TRUE)

  rosie_point = rosie$beta[upper.tri(rosie$beta)]
  rosie_se = sqrt(diag(solve(rosie$opt$hessian)))[-(1:sum(species_is_variable))]

  ######## BayesComm and partial BayesComm ########
  bc = BC(Y = x, model = "community", its = 1000)

  bc_pcors = sapply(
    1:nrow(bc$trace$R), 
    function(i){
      Sigma = matrix(0, nrow = ncol(x), ncol = ncol(x))
      Sigma[upper.tri(Sigma)] <- bc$trace$R[i, ]  # Fill in upper triangle
      Sigma <- Sigma + t(Sigma)                   # Fill in lower triangle
      diag(Sigma) <- 1  # Diagonal equals 1 in multivariate probit model
      pcor = cor2pcor(Sigma)
      pcor[upper.tri(pcor)]
    }
  )

  bind_rows(
    data_frame(
      rep_name = rep_name,
      n_sites = n_sites,
      sp1 = sp1,
      sp2 = sp2,
      method = "Markov network",
      truth = truth,
      estimate = rosie_point, 
      lower = qnorm(.025, rosie_point, rosie_se), 
      upper = qnorm(.975, rosie_point, rosie_se)
    ),
    data_frame(
      rep_name = rep_name,
      n_sites = n_sites,
      sp1 = sp1,
      sp2 = sp2,
      method = "GLM",
      truth = truth,
      estimate = coef_matrix[upper.tri(coef_matrix)], 
      lower = NA, 
      upper = NA
    ),
    data_frame(
      rep_name = rep_name,
      n_sites = n_sites,
      sp1 = sp1,
      sp2 = sp2,
      method = "correlation",
      truth = truth,
      estimate = corr[upper.tri(corr)], 
      lower = NA, 
      upper = NA
    ),
    data_frame(
      rep_name = rep_name,
      n_sites = n_sites,
      sp1 = sp1,
      sp2 = sp2,
      method = "partial correlation",
      truth = truth,
      estimate = p_corr[upper.tri(p_corr)], 
      lower = NA, 
      upper = NA
    ),
    data_frame(
      rep_name = rep_name,
      n_sites = n_sites,
      sp1 = sp1,
      sp2 = sp2,
      method = "partial BayesComm",
      truth = truth,
      estimate = rowMeans(bc_pcors),
      lower = apply(bc_pcors, 1, quantile, .025),
      upper = apply(bc_pcors, 1, quantile, .975)
    ),
    data_frame(
      rep_name = rep_name,
      n_sites = n_sites,
      sp1 = sp1,
      sp2 = sp2,
      method = "BayesComm",
      truth = truth,
      estimate = colMeans(bc$trace$R),
      lower = apply(bc$trace$R, 2, quantile, .025),
      upper = apply(bc$trace$R, 2, quantile, .975)
    )
  )
}
```

Run the above function on all the files:

```{r, cache = FALSE, eval = TRUE}

# Find all the csv files in the fakedata/matrices folder,
# then drop .csv
identifiers = dir("fakedata/matrices", pattern = "\\.csv$") %>%
  gsub("\\.csv$", "", .)

mclapply(identifiers, fit_all, mc.cores = mc.cores, mc.preschedule = FALSE, mc.silent = TRUE) %>% 
  bind_rows() %>% 
  write.csv(file = "estimates.csv")
```