### 2015-11-02 Harris_2015/manuscript-materials/draft.Rmd
etc

\chead{\ifthenelse{\value{page}=1}{}{\textit{Species interactions in Markov networks}}}
\renewcommand{\headrulewidth}{0pt}

\setlength{\parskip}{1pt}

**Title:** Inferring species interactions from co-occurrence data with Markov
networks

**Author:** David J. Harris: Population Biology; 1 Shields Avenue, Davis CA, 95616

**Abstract:** Inferring species interactions from co-occurrence data is one of
the most controversial tasks in community ecology. One difficulty is that a
single pairwise interaction can ripple through an ecological network and produce
surprising indirect consequences. For example, the negative correlation between
two competing species can be reversed in the presence of a third species that is
capable of outcompeting both of them. Here, I apply models from statistical
physics, called Markov networks or Markov random fields, that can predict
the direct and indirect consequences of any possible species interaction matrix.
Interactions in these models can also be estimated from observed co-occurrence
rates via maximum likelihood, controlling for indirect effects. Using simulated
landscapes with known pairwise interaction strengths, I evaluated Markov
networks and six existing approaches. The Markov networks consistently
outperformed other methods, correctly isolating direct interactions between
species pairs even when indirect interactions or abiotic factors largely
overpowered them. Two computationally efficient approximations, based on
controlling for indirect effects with linear or generalized linear models, also
performed well. Indirect effects reliably caused a common null modeling approach
to produce incorrect inferences, however.

**Key words:** Ecological interactions; Occurrence data; Species associations;
Markov network; Markov random field; Ising model; Biogeography;
Presence–absence matrix; Null model

\noindent\textbf{Introduction}

\setlength{\parindent}{2em}

\noindent
To the extent that nontrophic species interactions (such as competition) affect
community assembly, ecologists might expect to find signatures of these
interactions in species composition data [@macarthur_population_1958;
@diamond_island_1975]. Despite decades of work and several major controversies,
however [@lewin_santa_1983; @strong_ecological_1984; @connor_checkered_2013],
existing methods for detecting competition's effects on community structure are
unreliable [@gotelli_empirical_2009]. In particular, species' effects on one
another can become lost in a web of indirect effects. For example, the
competitive interaction between the two shrub species in Figure 1A is
obscured by their shared tendency to occur in unshaded areas (Figure 1B). While
ecologists have long known that indirect effects can overwhelm direct ones at
the landscape level [@dodson_complementary_1970; @levine_competitive_1976], the
vast majority of our methods for drawing inferences from observational data do
not control for these effects [e.g. @diamond_island_1975;
@strong_ecological_1984; @gotelli_empirical_2009; @veech_probabilistic_2013;
@pollock_understanding_2014]. To the extent that indirect interactions like
those in Figure 1 are generally important [@dodson_complementary_1970], existing
methods will not provide much evidence regarding species interactions.

While competition doesn't reliably reduce co-occurrence rates at the
whole-landscape level (as most methods assume), it nevertheless does leave a
signal in the data (Figure 1C). For example, after partitioning the data set
into shaded and unshaded sites, there will be co-occurrence deficits in
each subset that wouldn't otherwise be apparent. More generally, controlling for
other species in the network will often be important for obtaining reliable
estimates of direct (conditional, or all-else-equal) effects.  This kind of
precision is difficult to obtain from null models, which only simulate the most
extreme possible scenario: that *all* direct and indirect interactions are
exactly zero. Nevertheless, null models have dominated this field for more than
three decades [@strong_ecological_1984; @gotelli_empirical_2009].

Following @azaele_inferring_2010, this paper shows that Markov networks
[undirected graphical models also known as Markov random fields;
@murphy_machine_2012] can provide a framework for understanding the
landscape-level consequences of pairwise species interactions,
and for estimating them from observed presence-absence matrices. Markov networks
have been used in many scientific fields in similar contexts for decades, from
physics [where nearby particles interact magnetically; @cipra_introduction_1987]
to spatial statistics [where adjacent grid cells have correlated values;
@harris_contact_1974; @gelfand_modelling_2005]. While community ecologists
explored some related approaches in the 1980's [@whittam_species_1981],
they used severe approximations that led to unintelligible results
[e.g. "probabilities" greater than one; @gilpin_factors_1982].

Below, I demonstrate Markov networks' ability to produce exact predictions
about the direct and indirect consequences of an interaction matrix, and also to
make inferences about the species interactions that contributed to an observed
set of co-occurrences. Using simulated data sets where the "true" interactions
are known, I compare this approach with several existing methods. Finally, I
discuss opportunities for extending the approach presented here to other
problems in community ecology, e.g. quantifying the overall effect of species
interactions on occurrence rates [@roughgarden_competition_1983] and
disentangling the effects of biotic versus abiotic interactions on species
composition [@pollock_understanding_2014].

\noindent\textbf{Methods}

\noindent
\textbf{Markov networks.}
Markov networks provide a framework for translating back and forth between the
conditional (all-else-equal) relationships among species (Figure 1C) and the
kinds of species assemblages that these relationships produce. Here, I show how
a set of conditional relationships can be used to determine how groups of species
can co-occur. Methods for estimating conditional relationships from data are
discussed in the next section.

A Markov network defines the relative probability of observing a given vector of
species-level presences (1s) and absences (0s), $\vec{y}$, as

\centering

$\displaystyle{p(\vec{y}; \alpha, \beta) \propto exp(\sum_{i}\alpha_i y_i + \sum_{<ij>}\beta_{ij}y_i y_j),}$

\raggedright
\setlength{\parindent}{2em}

\noindent where the second sum is over all $\frac{1}{2}n(n-1)$ pairs of $n$
species. In this model, $\alpha_{i}$ is an intercept term determining the amount
that the presence of species $i$ contributes to the log-probability of
$\vec{y}$; it directly controls the prevalence of species $i$. Similarly,
$\beta_{ij}$ is the amount that the co-occurrence of species $i$ and species $j$
contributes to the log-probability; it controls the conditional relationship
between two species, i.e. the probability that they will be found together,
after controlling for the other species in the network (Figure 2A, Figure 2B).
For example, if $\beta_{ij} = +2$, then each species' odds of occurrence would be
$e^2$ times higher when the other one is present (as compared with otherwise
equivalent sites). Because the relative probability of a presence-absence vector
increases when positively-associated species co-occur and decreases when
negatively-associated species co-occur, the model tends---all else equal---to
produce assemblages that have many pairs of positively-associated species and
relatively few pairs of negatively-associated species (exactly as an ecologist
might expect). Of course, if all else is *not* equal (e.g. Figure 1, where the
presence of one competitor is associated with release from another competitor),
then predicting species' overall co-occurrence rates can be more complicated,
and may require summing over the different possible assemblages (Figure 2B).

\noindent \textbf{Estimating $\alpha$ and $\beta$ coefficients from presence-absence data.}
In the previous section, the values of $\alpha$ and $\beta$ were known and the
goal was to make predictions about possible species assemblages. In
practice, however, ecologists will often need to estimate the parameters from
an observed co-occurrence matrix (i.e. from a set of independent $\vec{y}$ vectors
indicating which species are present at each site on the landscape). When the number of species
is reasonably small, one can compute exact maximum likelihood estimates for all
of the $\alpha$ and $\beta$ coefficients given a presence-absence matrix by
optimizing $p(\vec{y}; \alpha, \beta)$. Fully-observed Markov networks like the
ones considered here have unimodal likelihood surfaces [@murphy_machine_2012],
ensuring that this procedure will always converge on the global maximum. This
maximum represents the unique combination of $\alpha$ and $\beta$ coefficients
that would be expected to produce exactly the observed co-occurrence frequencies
on average [i.e., maximizing the likelihood ensures that the sufficient
statistics of the model distribution match the sufficient statistics of the data;
@murphy_machine_2012]. I used the rosalia package [@harris_rosalia_2015] for the
R programming language [@r_core_team_r_2015] to optimize the Markov network
parameters.

\noindent \textbf{Simulated landscapes.}
I simulated several sets of landscapes using known parameters to evaluate
different statistical methods' performance. The first set of landscapes included
the three competing species shown in Figure 1. For each of 1000 replicates, I
generated a landscape with 100 sites by sampling from a probability distribution
defined by the figure's interaction coefficients (Appendix 1). Each of the
methods described below was then evaluated on its ability to correctly infer
that the two shrub species competed with one another, despite their frequent
co-occurrence.

I then generated landscapes with up to 20 interacting species at 25, 200, or
1600 sites using three increasingly complex models (50 replicates for each
combination of size and model). As described in Appendix [[X]], I randomly drew
the “true” coefficient values for each replicate landscape
so that most species pairs interacted negligibly, a few pairs interacted very
strongly, and competition was three times more common than facilitation.

The first set of 20-species landscapes, like the landscapes with three species,
were generated directly from a Markov network to ensure that the model could
recover the parameters used to generate the "observed" co-occurrence data.
Then, I added two environmental factors that varied from location to location
across the simulated landscapes, and simulated a new set of co-occurrence data
so that species' $\alpha$ coefficients depended on the local environment.
The latter set of simulated landscapes provide an important test of the methods’
ability to distinguish co-occurrence patterns that were generated from pairwise
biotic interactions from those that were generated by external forces like
abiotic environmental filtering. This task was made especially difficult
because---as with most analyses of presence-absence data for co-occurrence
patterns---the inference procedure did not have access to any information about
the environmental or spatial variables that helped shape the landscape [cf
@connor_checkered_2013].  I generated the final set of landscapes by simulating
the population dynamics of 20 interacting species, as described in Appendix [[X]].

\noindent \textbf{Recovering species interactions from simulated data.}
I compared seven techniques for determining the sign and strength of the
associations between pairs of species from simulated data (Appendix 3).
First, I used the rosalia package [@harris_rosalia_2015] to fit Markov network
models,  as described above. For the analyses with 20 species, I added a weak
regularizer (equivalent to a logistic prior with scale 2) to ensure that the
model estimates were always finite [[Appendix]].

I also evaluated six alternative methods: five from the existing literature,
plus a novel combination of two of these methods. The first alternative interaction
metric was the sample correlation between species' presence-absence vectors,
which summarizes their marginal association.  Next, I used partial correlations,
which summarize species' conditional relationships. This approach is
common in molecular biology [@friedman_sparse_2008], but is rare in ecology
(see @albrecht_spatial_2001 and @faisal_inferring_2010 for two exceptions).
In the context of non-Gaussian data, the partial correlation can be thought of
as a computationally efficient approximation to the full Markov network model
[@loh_structure_2013]. Because partial correlations are undefined for
landscapes with perfectly-correlated species pairs, I used a regularized
estimate based on James-Stein shrinkage, as implemented in the corpcor package’s
`pcor.shrink` function with the default settings [@schafer_corpcor_2014].

The third alternative, generalized linear models (GLMs), also provide a
computationally efficient approximation to the Markov network
[@lee_learning_2012].  Following @faisal_inferring_2010,  I fit regularized
logistic regression models [@gelman_weakly_2008] for each species, using the
other species on the landscape as predictors. This produced two interaction
estimates for each species pair; one for the effect of species $i$ on species
$j$ and one for the reverse. These two coefficients are not identifiable from
the data, however [@schmidt_modeling_2012], so I used their average as an
overall measure of the overall relationship between species $i$ and species $j$.

The next method, described in @gotelli_empirical_2009, involved
simulating new landscapes from a null model that retains the row and column
sums of the original matrix [@strong_ecological_1984]. I used the $Z$-scores
computed by the Pairs software described in @gotelli_empirical_2009 as my null
model-based estimator of species interactions.

The last two estimators used the latent correlation matrix estimated by the
BayesComm package [@golding_bayescomm_2015] in order to evaluate the recent
claim that the correlation coefficients estimated by "joint species distribution
models" provide an accurate assessment of species’ pairwise interactions
[@pollock_understanding_2014; see also @harris_generating_2015]. In addition to
using the posterior mean correlation [@pollock_understanding_2014], I also used
the posterior mean *partial* correlation, which might be able to control for
indirect effects.

\noindent \textbf{Evaluating model performance.}
For the simulated landscapes based on Figure 1, I assessed whether each method's
test statistic indicated a positive or negative relationship between the two
shrubs (Appendix 1). For the null model (Pairs), I calculated statistical
significance using its $Z$-score. For the Markov network, I used the Hessian
matrix to generate approximate confidence intervals.

I then evaluated the relationship between each method's estimates and the "true"
interaction strengths among all of the species pairs from the larger simulated
landscapes. This determined which of the methods provide a consistent way to
know how strong species interactions are---regardless of which species were
present in a particular data set or how many observations were taken. To ensure
that the different test statistics (e.g. correlations versus $Z$ scores) were on
a common scale, I rescaled them using linear regression through the origin. I
then calculated the proportion of variance explained for different combinations
of model type and landscape size (compared with a null model that assumed all
interaction strengths to be zero).

\noindent
\textbf{Results}

\noindent \textbf{Three species.}
As shown in Figure 1, the marginal relationship between the two shrub species
was positive---despite their competition for space at a mechanistic level---due
to indirect effects of the dominant tree species. As a result, the correlation
between these species was positive in 94% of replicates, and the
randomization-based null model falsely reported positive associations 100% of
the time. Worse, more than 98% of these false conclusions were statistically
significant. The partial correlation and Markov network estimates, on the other
hand, each correctly isolated the direct negative interaction between the shrubs
from their positive indirect interaction 94% of the time (although the
confidence intervals overlapped zero in most replicates).

\noindent
\textbf{Twenty species.}
Despite some variability across contexts (Figure 3), the four methods that
controlled for indirect effects (the Markov network, the generalized linear
models, and the two partial correlation-based methods) outperformed those
that did not.  Of these, the Markov network consistently performed the best.
As anticipated by @lee_learning_2012, generalized linear models closely
approximated the Markov network estimates, especially when the data sets were
very large. As reviewed in @gotelli_empirical_2009, however, most analyses in
this field involve fewer than 50 sites, so model performance on the smaller
landscapes where the gap was larger is most important. As shown in Appendix
[[X]], the standard errors associated with the estimates in Figure 3 are small
(less than 0.01), so the differences among methods should not be attributed to
sampling error.

Of the methods that did not control for indirect effects, Figure 3 shows that
simple correlation coefficients provided a more reliable indicator of species'
true interaction strengths than either the joint species distribution model
(BayesComm) or the null model (Pairs). The estimates from these approaches
were tightly correlated (after controlling for the size of the landscape)
suggesting that the null model only contains a noisy version of the
same information that could be obtained more easily and interpretably with
simple correlation coefficients (Figure 4B).

Finally, we can evaluate the models' statistical inferences (focusing on the
first two simulation types, for which the true interaction rates are easiest to
interpret).  The Markov network's approximate Type I error rate (defined here as
its probability of rejecting the null hypothesis for a pair of species where
$|\beta_{ij}|<0.1$) depended on the simulation type: 0.03 for simulations that
matched the model's assumptions, versus 0.15 for simulations that included
environmental heterogeneity [[Appendix]].  For the null model, these rates were
0.31 and 0.50, respectively---far higher than the nominal 0.05 rate. Figure 4C
shows, across a range of true interaction strengths, the probability that the
null model or the Markov network will predict the wrong sign of the interaction
with 95% confidence.  The null model makes such errors nearly 8 times as often
as the Markov network, even though it only reject the null hypothesis twice as
often overall [[Appendix]].  The Markov network's errors were also more
concentrated around 0, so it never misclassified strong interactions like the
null model did.

\noindent
\textbf{Discussion}

\noindent
The results presented above show that Markov networks can reliably recover
species' pairwise interactions from observational data, even for cases where
environmental heterogeneity and indirect interactions cause ecologists' typical
null modeling approaches to reliably fail. Partial covariances and generalized
linear models can both provide computationally efficient approximations, but
with somewhat lower accuracy, especially for typically-sized data sets with small
numbers of sites [@gotelli_empirical_2009]. The difference in accuracy may be
larger for real data sets than for the simulated landscapes in Figure 3,
however; linear approximations to the Markov network make larger errors when
the interaction matrix is structured [e.g. due to guilds or trophic levels;
@loh_structure_2013]. Similarly, the separate generalized linear models
for each species can severely overfit in some cases [@lee_learning_2012]. The
full Markov network should thus be preferred to the approximations when
it is computationally tractable.

Compositional data only contains enough degrees of freedom to estimate one
interaction per species pair [@schmidt_modeling_2012], so none of these methods
can identify the exact nature of the pairwise interactions (e.g. which species
in a positively-associated pair is facilitating the other). To estimate
asymmetric interactions, such as commensalism or predation, ecologists could use
time series, behavioral observations, manipulative experiments, or natural
history. These other sources of information could also be used to augment the
likelihood function with an informative prior distribution, reducing ecologists'
error and uncertainty relative to Figure 3's results.

Despite their limitations, Markov networks have enormous potential to improve
ecological understanding. In particular, they make many fewer errors than existing
approaches, and can make precise statements about the conditions where indirect
interactions will overwhelm direct ones. They also provide a simple answer to
the question of how competition should affect a species' overall prevalence,
which has important implications for community-level modeling
[@strong_ecological_1984]. Specifically, Equation 1 can be used to calculate the
expected prevalence of a species in the absence of biotic influences as
$e^\alpha/(e^{0} + e^\alpha)$. Competition's effect on prevalence can then be
estimated by comparing this value with the observed prevalence (e.g. comparing
Figure 2D with Figure 2C). This novel quantitative result conflicts with most of
our null models, which unreasonably assume that prevalence would be the exactly
same in the absence of competition as it is in the observed data
[@roughgarden_competition_1983].

Markov networks---particularly the Ising model for binary networks---are very
well understood, having been studied for nearly a century
[@cipra_introduction_1987]. Tapping into this framework would thus allow
ecologists to take advantage of into a vast set of existing discoveries and
techniques for dealing with indirect effects, stability, and alternative stable
states. Numerous extensions to the basic network are possible as well. For
example, the states of the interaction network can be modeled as a function of
the local abiotic environment [@lee_learning_2012], which would help incorporate
networks of biotic interactions into species distribution models
[@pollock_understanding_2014] and lead to a better understanding of the
interplay between biotic and abiotic effects on community structure.
Alternatively, models could allow one species to alter the relationship between
two other species [@tjelmeland_markov_1998; cf @bruno_inclusion_2003].

Finally, the results presented here have important implications for
ecologists' continued use of null models for studying species interactions. Null
and neutral models can be useful for clarifying our thinking
[@harris_occupancy_2011; @xiao_strong_2015], but deviations from a particular
null model must be interpreted with care [@roughgarden_competition_1983]. Even
in small networks with three species, it may simply not be possible to implicate
specific ecological processes like competition by rejecting a general-purpose
null [@gotelli_empirical_2009], especially when the test statistic is
effectively just a correlation coefficient (Figure 4B). When the non-null
backdrop is not controlled for, Type I error rates can skyrocket, the apparent
sign of the interaction can change, and null models can routinely produce
misleading inferences (Figure 1, Figure 4C, @gotelli_empirical_2009).

Controlling for indirect effects via simultaneous estimation of multiple
ecological parameters seems like a much more promising approach: to the extent
that the models' relative performance on real data sets is similar to the range
of results shown in Figure 3, scientists in this field could more than double
their explanatory power by switching from null models to Markov networks (or
increase it nearly as much with linear or generalized linear approximations).
Regardless of the methods ecologists ultimately choose, controlling for indirect
effects could clearly improve our understanding of species' direct effects on
one another and on community structure.

\noindent \textbf{Acknowledgements:}
This work benefited greatly from discussions with A. Sih, M. L. Baskett, R.
McElreath, R. J. Hijmans, A. C. Perry, and C. S. Tysor. Additionally, A. K.
Barner, E. Baldridge, E. P. White, D. Li, D. L. Miller, N. Golding, N. J.
Gotelli, C. F. Dormann, and two anonymous reviewers provided very helpful
feedback on the text. This research was partially supported by a Graduate
Research Fellowship from the US National Science Foundation and by the Gordon
and Betty Moore Foundation's Data-Driven Discovery Initiative through Grant
GBMF4563 to E. P. White.

\setlength{\parindent}{0cm}

\noindent \textbf{Figure 1: A.} A small network of three competing species.
The tree (top) tends not to co-occur with either of the two shrub species, as
indicated by the strongly negative coefficient linking them. The two shrub
species also compete with one another, but more weakly (circled coefficient).
**B.** In spite of the competitive interactions between the two shrub species,
their shared tendency to occur in locations without trees makes their occurrence
vectors positively correlated (circled). **C.** Controlling for trees with a
conditional (all-else-equal) approach such as a partial covariance or a Markov
network leads to correct identification of the negative shrub-shrub interaction
(circled).

\noindent \textbf{Figure 2: A.} A small Markov network, defined by its $\alpha$ and
$\beta$ values.  The abiotic environment favors the occurrence of each species
($\alpha >0$), particularly species 2 ($\alpha_2 > \alpha_1$). The negative
$\beta_{12}$ coefficient is consistent with competition between the two species.
**B.** The coefficients determine the probabilities of all four possible
presence-absence combinations for Species 1 and Species 2. $\alpha_1$ is added
to the exponent whenever Species 1 is present ($y_1 = 1$), but not when it is
absent ($y_1 = 0$).  Similarly, the exponent includes $\alpha_2$ only when
species $2$ is present ($y_2 = 1$), and includes $\beta_{12}$ only when both are
present ($y_1y_2 = 1$). The normalizing constant $Z$, ensures that the four
probabilities sum to 1.  In this case, $Z$ is about 18.5.  **C.** The expected
frequencies of all possible co-occurrence patterns between the two species of
interest, as calculated in the previous panel. **D.** Without competition (i.e.
with $\beta_{12}=0$, each species would occur more often.

\noindent \textbf{Figure 3:} Proportion of variance in interaction
coefficients explained by each method versus number of sampled locations across
the three simulation types.

\noindent \textbf{Figure 4: A.} The Markov network's estimated interaction
coefficients were generally very similar to the GLM estimates. **B.** The null
model's estimates typically matched the (negative) correlation coefficient,
after controlling for landscape size. **C.** For any given interaction strength,
the null model was much more likely to misclassify its sign with 95% confidence
than the Markov network was.

\noindent \textbf{References:}

\setlength{\parindent}{-1em}
\setlength{\leftskip}{1em}