### 2015-11-05 Harris_2015/manuscript-materials/response-to-editor.md
add fakedata and updates

Dear Mr. Harris,

Thank you very much for submitting your revised manuscript "Estimating species interactions from observational data with Markov networks" (Statistical Report) for review by Ecology. Many aspects of the presentation are improved. Unfortunately my assessment is that the new community dynamics simulations turned out to be a step in the wrong direction, making it harder rather than easier to assess the new method and compare it to older ones. I can see that this new material resulted from taking Reviewer #1's suggestion to heart, but the bottom line is that it didn't turn out well.

First, I don't think you needed to remove the extensive set of potentially very informative simulations that you had done previously but, if anything, add to them. (I know the 20-page Report limit is tight, but you have no limits in the supplement.) Evaluation of new methods when assumptions are not met builds naturally on first having evaluated the method when its assumptions are met, not skipping over that step. However, you have removed the lion's share of such comparisons, leaving only the simple 3 species case before diving into the community dynamics simulations. I was left unable to know how the method performs on its own terms before being faced with evaluation of how it performs on very complicated terms.

> I appreciate this perspective.  I felt very ambivalent about including these simulations in the first place, and had many of the same concerns that you raise.  I ultimately decided to include them because a substantial number of colleagues that read the preprint had similar concerns to Reviewer 1: they wanted to see how the different methods would perform on more realistic data sets and especially on data sets that were not simulated from a model that resembled a Markov network.  I do agree, however, that it is important to show that the model performs well on its own terms before diving into a more complicated model, and that omitting the simpler simulations made the paper worse. For this reason, I now include three different simulation regimes for the 20-species landscapes: 1. Simulating directly from a Markov network, 2. Simulating from a Markov network with spatial heterogeneity in the α parameters (sometimes called a conditional random field), and 3. Simulating with a dynamic model.

> For this revision, I considered a number of alternative simulation methods for the dynamic model that could address the concerns of Reviewer 1 and my colleagues, but they all had serious problems.  Using a set of GLMs as the generative model (as you suggest below) would either involve directed cycles (if β_ij≠β_ji) or would match the Markov network exactly (if β_ij=β_ji). Allowing directed cycles in the generative model would cause a number of problems. Most importantly, these models would not be identifiable from the presence-absence vectors they generated, as an infinite number of different parameter settings could all generate a given joint distribution. On the other hand, re-labeling the simulation network as a set of GLMs with equality constraints on the coefficients might have alleviated some readers' concerns, but only by hiding the fact that it was actually the same model.

Your proposed community dynamics model would raise a host of concerns. You state it is a large population approximation but then include demographic stochasticity, which is typically of diminishing importance for large populations and outweighed by environmental stochasticity, which you omit.

> I now refer to the differential equation as the model's deterministic backbone, rather than as an approximation.

The choice of parameters to draw from is arbitrary and ultimately discernible only by reading code, and the mutualism term appears to be arbitrary.

> This is true, but it was also true of all the other models I could envision where mutualism was guaranteed not to cause unbounded population explosions regardless of the parameters chosen.

You may disagree, and one could debate all aspects, but that is really the point: one would have wanted a simpler and more standard model so comparison of the statistical methods is not tangled in these issues.

> The simpler models are now included.

I can see you made a serious effort, but the outcome is not something I think readers will gain value from. The ultimate problem is that we are left with no comprehension of how the data generated by these processes suit the assumptions of the various methods. Does the model actually generate data that are well approximated by the Markov network assumptions or not? Trying to understand that would be a theoretical ecology exercise that, for your proposed model, does make sense to include in a paper about the statistical methods.

> Based on the number of readers that have asked for something along these lines (including Reviewer 1), I think that there is some value.  The new Figure 3 shows that the addition of population dynamics and per-capita interactions does weaken the relationship between the simulation coefficients and all of the estimation methods, but that it does not change the rank ordering of the different methods, which is an important result.

I suggest instead staying closer to the statistical models and including some additional cases to probe the method's robustness to violations of its assumptions and/or cases where one of the other parametric models in the competition generates the data. Which kinds of potential violations of assumptions to consider is for you to decide. For example, perhaps the cases of interest for the Markov network model would be if there really are 3-way interactions, or asymmetric interactions, or heterogeneity among sites in the model parameters, that contribute to the data generating process but are ignored in the analysis model. Indeed, describing what would constitute violations would help the reader. Another good way to probe would be to generate data from one of the other parametric models, such as the GLM.

> I now do this with the first two simulation types: the first type matches the model's assumptions and the second one adds heterogeneity among sites.  I now show that this violation decreases the models' R-squared values and increases their Type I error rates.

Note that often violations of assumptions do not create bias in parameter estimates (although sometimes they do) but rather incorrect assessments of uncertainty (inaccurate confidence intervals and Type I error rates in hypothesis tests). You have referred in multiple places to having applied a bootstrap as well as calculating Wald intervals, but it appears to me you have reported almost no results about validity of confidence intervals or hypothesis tests from those (there are a few results for the 3-species case). It would be important to do so.

> I now report approximate Type I error rates

It appears to me that you did not understand my point that you are presenting results that are integrated over parameter distributions, which appears still to be the case. What I meant is the following. Often one would want to see how a statistical method performs as a function of specific parameter values. For example, one might vary the value of an interaction coefficient (beta) from negative (competition) to zero to positive (mutualism) and run a set of simulations for each value of beta. This would allow evaluation of statistical power and accuracy of confidence intervals and Type I error rate as well as RMSE, all as a function of true parameter values. It would also allow assessment of whether the method works better for some parameter scenarios than others, which is not uncommon. Instead you are drawing parameters from distributions and then presenting performance assessments in aggregate from those results, such as the average R-squared over all cases of a given network size. Generally, and clearly in your case, such metrics can be written as expectations (integrations) over the distribution of parameters you sampled from. You are right there is nothing Bayesian about it, but I pointed out a Bayesian may feel justified in doing it because many Bayesian results take the same format. It is a question of presenting aggregated versus disaggregated results. Unfortunately, presenting only aggregated results leaves the reader unable to dig deeper into the disaggregated results. The analogy to sampling from an archipelago is lovely but irrelevant. It appears to me that since you have all the simulations at hand, and since you can place more detailed results in the supplement, you should do so. I leave it to you to decide how to do that.

> Showing how some of the results change with the value of beta is a great idea, and I have added several analyses along these lines in Figure 4 and Appendix 5.  In particular, I show how CI coverage and the probability that the model will predict the wrong sign for an interaction vary as a function of beta, in addition to global estimates of R-squared and Type I error rates.  Unfortunately, I am dealing with a 210-dimensional system (20 alphas and 190 betas), which limits my options.  Holding 209 parameters constant at arbitrary values doesn't seem like the right answer, since the results would depend on which values were chosen.  Apart from the trivial case where all the coefficients except one are fixed at zero (in which case the indirect interactions that make this paper's methods necessary would not exist), there doesn't seem to be any particular line through the 210-dimensional space that deserves this kind of special attention. Likewise, there isn't a good way for completely free-form exploration of this large space.  For this reason, the best compromise in this revision seemed like marginalizing over all but one variable at a time for these analyses.

Some other minor notes

- I don't think you have stated that the betas are symmetric. It appears that beta_ij = beta_ji and hence each could be interpreted as half of the interaction strength. E.g. on line 80, it would be e^4 if I take the preceding model equation as correct and include both beta_ij and beta_ji. Please fix it up if that is not right.

> I've made this issue much clearer in the revision: see lines [[XXXX]].  It should now be unambiguous that the sum in the likelihood should only include each pair once, and that each pair must share a single value for $\beta_{ij}$.

- Both times that I have seen the title anew, the phrase "estimating species interactions from observational data" made me think it would be about time-series method, since that is a common kind of "observational data" from which people try to estimate interactions. I recommend adding a descriptive term like "species composition data" or "observational occurrence data" or "presence/absence data." You decide. This could enter the first Abstract sentence as well.

> This is another good point.  I've changed the title to focus on co-occurrence data.  Now that I've included inferential statistics in the paper and evaluated their frequentist properties, I've also changed the word "estimating" back to "inferring".

- Line 44 is not stated well. You could replace "begin with the assumption" by "test the hypothesis" or insert "that the hypothesis of interest is that [all pairwaise interactions...]" or some such change.

- Line 69: The phrase "how groups of species can co-occur" seems imprecise or incorrect. Maybe "to determine species occurence models"?

- Lines 86-87 are not a good lead-in to the point of the rest of this paragraph. It sounds like you are about to talk about species interactions, again, but really the point of the paragraph is that normalizing the probabilities involves a difficult summation.

- Lines 72 & 101. I was willing to overlook the fact that you didn't define the vector y explicitly in terms of its elements on line 72 because perhaps that is really obvious. However on line 101 you then state maximum likelihood as if it would be done from a single y vector (and a single likelihood term) rather than a matrix of observations. So now I think you need more explicit notation to clarify the arrangements of your variables.

> The revised manuscript now clarifies that each $\vec{y}$ vector represents
the different species that could occur at a single site and that the full
co-occurrence matrix is made of "a set of independent $\vec{y}$ vectors
indicating which species are present at each site on the landscape." This should
clarify the relationship betweeen the $\vec{y}$ vectors and the matrix as well
as the relationship between the single-site likelihood and the landscape-level
likelihood (independence among sites).

- Line 130-132. This statement is unclear and unconvincing. Are you saying that the remoteness of the upper range of the prior justifies that the prior has no influence? That would not be correct since it is an issue of prior weightings in the range of heavy posterior weight. The simplest way to show a prior has little influence is to try a couple of different ones.

- Lines 136-137. Bootstrapping is a method to estimate uncertainty of an estimation procedure. It is unclear what you mean by using it to validate if a procedure gives "stable" estimates.

> The bootstrapping results have now been removed and replaced with the confidence intervals.

- Line 154: This statement is not clear.

- Figure 2B: This is odd notation because, if I follow, you have chosen a null symbol (0 with a slash) to indicate the event y_i = 0 but have chosen the label y_i to indicate the event y_i = 1. I think you want P[00], P[10], etc. or P[y1=0, y2=0], P[y1=1, y2=0], etc, or something like that. The caption states "relative probability" where I think you want "probability."

> This is an excellent point. The figure has been fixed (along with the caption).

- Figures in general: would benefit from titles on each subfigure.

> Subfigures now have titles

It's too bad we are in this unusual situation that your effort to respond to a reviewer made the manuscript worse. I will let you make another revision to get back on course. What I need you to do are: (1) remove the community dynamics model, put back in extensive results from the model of interest, and take a new approach to probing the model more thoroughly by including some scenarios with clearly defined violations of assumptions and/or data generated from one of the alternative statistical models. You may decide what to put in the online material since I recognize you are trying to stick to the 20-page report limit. (2) Include meaningful results about accuracy of uncertainties. These should be at your fingertips since you can calculate confidence intervals for every simulation. (3) Provide some disaggregated results. (4) Show how the regression steps work in about a page or two of supplement text, not requiring code reading.