### 2015-11-25 Harris_2015/simulation.Rmd

---
output: pdf_document
title: "Appendix 2: Simulating 20-species landscapes"
subtitle: "Inferring species interactions from co-occurrence data with Markov networks"
author: David J. Harris
fontsize: 11pt
---

## Simulating landscapes from known coefficients

The following function draws random coefficients for a Markov network of a pre-specified size.

```{r, cache=TRUE}
make_coefficients = function(n_spp, p_neg, mean_alpha){
  # Exponential distribution has lots of mass near 0 but has
  # a long tail.
  true_beta_magnitudes = rexp(choose(n_spp, 2))

  # Multiply some proportion of the interactions
  # by -1
  b = true_beta_magnitudes * sample(
    c(-1, 1), 
    size = length(true_beta_magnitudes), 
    prob = c(p_neg, 1 - p_neg),
    replace = TRUE
  )

  # Species' intercepts are normally distributed
  a = rnorm(n_spp, mean_alpha)

  # Return the simulated values.
  # The rosalia function stores pairwise parameters in the upper
  # triangle of an n-by-n matrix and stores the species intercepts
  # along the diagonal, so these values are named accordingly.
  c(alpha  = a, beta = b)
}
```

I used the following R function to generate "true" parameters with the method above, 
then simulate a presence-absence landscape based on those parameters using Markov chain Monte Carlo, and finally save the results to a "fakedata" folder.

I used Gibbs sampling as my Markov chain Monte Carlo transition operator.
In each round of Gibbs sampling, I cycled through all the species, randomly
updating each one's presence/absence vector in response to its conditional
occurrence probability:

$$p(y_i) = \mathrm{logistic}(\alpha_i + \sum_j\beta_{ij}y_j),$$

where the logistic function is $\frac{1}{1+e^{-x}}$.

For the abundance-based simulations, abundance was simulated with a Poisson
distribution with rate parameter given by 
$\mathrm{log}(1 + exp(\alpha_i + \sum_j\beta_{ij}y_j))$.

```{r, cache = TRUE}
simulate_data = function(n_spp, n_sites, rep_name, n_gibbs, n_env, sd, f, rdist, p_neg, mean_alpha){
  # n_spp:     number of species to include in the landscape
  # n_sites:   number of sites to include in the landscape
  # rep_name:  an identifier to use for the landscape replicate
  # n_gibbs:   number of Gibbs samples to perform
  # n_env:     number of environmental variables to simulate
  # sd:        standard deviation of environmental variables (can be zero)
  # f:         inverse link function (see above for two examples)
  # rdist:     a function for sampling a random value from a distribution
  # p_neg:     proportion of negative interactions (e.g. competition)
  # mean_alpha: the intercept value for the average species

  # Determine the "true" parameters for the simulated assemblage
  par = make_coefficients(n_spp, p_neg, mean_alpha)

  # "True" interaction strengths, to save for later
  truth = par[-(1:n_spp)]

  # "True" intercepts, possibly adjusted below by environment
  alpha = par[1:n_spp]

  # Turn the interaction values into an n-by-n matrix
  # Start with empty matrix; fill in upper triangle; 
  # then fill in lower triangle with its transpose
  beta = matrix(0, n_spp, n_spp)
  beta[upper.tri(beta)] = truth
  beta = beta + t(beta)

  # Environmental states are normally distributed
  env = matrix(rnorm(n_sites * n_env), ncol = n_env)

  alpha_env = matrix(rnorm(n_spp * n_env, sd = sd), nrow = n_env)

  # Simulate the landscape from known process with Gibbs sampling

  # Landscape starts as if betas were all zero. Each species' occurrence probability
  # or abundance depends on its alpha value and on the environment (assuming alpha_env 
  # is not zero).
  x = matrix(
    f(rep(1, n_sites) %*% t(alpha) + env %*% alpha_env), 
    nrow = n_sites, 
    ncol = n_spp
  )

  # Gibbs sampling
  for(i in 1:n_gibbs){

    # Each round of Gibbs sampling updates one species (column) across all sites
    # according to its conditional probability (i.e. conditional on environment
    # and the other species that are present).
    for(j in 1:n_spp){
      x[,j] = rdist(
        nrow(x),
        f(x %*% beta[ , j] + alpha[j] + env %*% alpha_env[,j])
      )
    }
  }

  # Collapse abundance data to presence/absence and store
  # it as integer values rather than true/false
  x = x > 0
  mode(x) = "integer"

  colnames(x) = paste0("V", 1:n_spp)

  # Save the results in a "fake data" folder

  file_stem = paste(n_sites, rep_name, sep = "-")

  # Gotelli and Ulrich's Pairs software rejects empty sites, so I remove them here
  x_subset = x[rowSums(x) != 0, colSums(x) != 0]

  # Save the matrix of presence/absence observations
  write.csv(
    x, 
    file = paste0("fakedata/matrices/", file_stem, ".csv")
  )

  # Gotelli and Ulrich's Pairs method expects the data matrices to be transposed,
  # So I save them separately
  write.table(
    t(x_subset), 
    file = paste0("fakedata/matrices/", file_stem, "-transposed.txt"), 
    quote = FALSE
  )

  # Save the "true" species interactions
  write(
    truth, 
    file = paste0("fakedata/truths/", file_stem, ".txt"), 
    ncolumns = length(truth)
  )
}
```

```{r}
# Define a convenience function for Bernoulli random samples
rbern = function(n, prob){
  rbinom(n = n, size = 1, prob = prob)
}
```

Loop to simulate and save all the landscapes:

```{r}
set.seed(1)
n_spp = 20

for(n_sites in c(25, 200, 1600)){
  for(type in c("no_env", "env", "abund")){
    for(i in 1:50){
      simulate_data(
        n_spp = n_spp, 
        n_sites = n_sites, 
        n_gibbs = 1000,
        n_env = 2,
        rep_name = paste0(type, i),
        sd = ifelse(type == "env", 2, 0),
        f = if(type == "abund"){function(x){log(1 + exp(x))}}else{plogis},
        rdist = if(type == "abund"){rpois}else{rbern},
        p_neg = if(type == "abund"){1}else{0.75},
        mean_alpha = if(type == "abund"){5}else{-1}
      )
    }
  }
}

```