### 2016-02-11 Harris_2015/manuscript-materials/draft.Rmd
editor comments

---
output: pdf_document
---
\chead{\ifthenelse{\value{page}=1}{}{\textit{Inferring species interactions}}}
\renewcommand{\headrulewidth}{0pt}

\setlength{\parskip}{1pt}

**Title:** Inferring species interactions from co-occurrence data with Markov
networks

**Author:** David J. Harris: Population Biology; 1 Shields Avenue, Davis CA, 95616

**Abstract:** Inferring species interactions from co-occurrence data is one of
the most controversial tasks in community ecology. One difficulty is that a
single pairwise interaction can ripple through an ecological network and produce
surprising indirect consequences. For example, the negative correlation between
two competing species can be reversed in the presence of a third species that is
capable of outcompeting both of them. Here, I apply models from statistical
physics, called Markov networks or Markov random fields, that can predict
the direct and indirect consequences of any possible species interaction matrix.
Interactions in these models can be estimated from observed co-occurrence
rates via maximum likelihood, controlling for indirect effects. Using simulated
landscapes with known pairwise interaction strengths, I evaluated Markov
networks and six existing approaches. The Markov networks consistently
outperformed other methods, correctly isolating direct interactions between
species pairs even when indirect interactions or abiotic factors largely
overpowered them. Two computationally efficient approximations, based on
controlling for indirect effects with linear or generalized linear models, also
performed well. Indirect effects reliably caused a common null modeling approach
to produce incorrect inferences, however.

**Key words:** Ecological interactions; Occurrence data; Species associations;
Markov network; Markov random field; Ising model; Biogeography;
Presence–absence matrix; Null model

\noindent\textbf{Introduction}

\setlength{\parindent}{2em}

\noindent
To the extent that nontrophic species interactions (such as competition) affect
community assembly, ecologists might expect to find signatures of these
interactions in species composition data [@macarthur_population_1958;
@diamond_island_1975]. Despite decades of work and several major controversies,
however [@lewin_santa_1983; @strong_ecological_1984; @connor_checkered_2013],
existing methods for detecting competition's effects on community structure are
unreliable [@gotelli_empirical_2009]. In particular, species' effects on one
another can become lost in a web of indirect effects. For example, the
competitive interaction between the two shrub species in Figure 1A is
obscured by their shared tendency to occur in unshaded areas (Figure 1B). While
ecologists have long known that indirect effects can overwhelm direct ones 
[@levine_competitive_1976], most methods for drawing inferences from 
co-occurrence data do not control for these effects [e.g. @diamond_island_1975;
@strong_ecological_1984; @gotelli_empirical_2009; @veech_probabilistic_2013;
@pollock_understanding_2014]. As a result, ecologists do not have tools that allow
them to isolate direct interactions from indirect effects.

While competition doesn't reliably reduce co-occurrence rates at the
whole-landscape level (as most methods assume), it does still leave a
signal in the data (Figure 1C). After controlling for the presence
of the tree species (e.g. by splitting the data set into shaded and unshaded 
sites or by using a model that estimates conditional relationships amongs 
species), the two shrubs do have a negative association, with fewer co-occurrences 
in unshaded areas than would be expected if they occurred independently of one
another.

Following @azaele_inferring_2010, this paper shows that Markov networks
[undirected graphical models also known as Markov random fields;
@murphy_machine_2012] can provide a framework for understanding the
landscape-level consequences of pairwise species interactions,
and for estimating them from observed presence-absence matrices. Markov networks
have been used in diverse scientific fields in similar contexts for decades, from
physics [where nearby particles interact magnetically; @cipra_introduction_1987]
to spatial statistics [where adjacent grid cells have correlated values;
@harris_contact_1974; @gelfand_modelling_2005]. While community ecologists
explored some related approaches in the 1980's [@whittam_species_1981],
they used severe approximations that led to unintelligible results
[e.g. "probabilities" greater than one; @gilpin_factors_1982].

Below, I demonstrate Markov networks' ability to produce exact predictions
about the direct and indirect consequences of an interaction matrix, and also to
make inferences about the interaction matrix based on co-occurrence rates. Using 
simulated data sets where the "true" interactions are known, I compare this 
approach with several existing methods. Finally, I
discuss opportunities for extending the approach presented here to other
problems in community ecology, e.g. quantifying the overall effect of species
interactions on occurrence rates [@roughgarden_competition_1983] and
disentangling the effects of biotic versus abiotic interactions on species
composition [@pollock_understanding_2014].

\noindent\textbf{Methods}

\noindent
\textbf{Markov networks.}
Markov networks provide a framework for translating back and forth between the
conditional (all-else-equal) relationships among species (Figure 1C) and the
kinds of species assemblages that these relationships produce. Here, I show how
a set of conditional relationships can determine species composition. Methods for
estimating conditional relationships from data are discussed in the next section.

A Markov network defines the relative probability of observing a given vector of
species-level presences (1s) and absences (0s), $\vec{y}$ at a site, as

\centering

$\displaystyle{p(\vec{y}; \alpha, \beta) \propto exp(\sum_{i}\alpha_i y_i + \sum_{<ij>}\beta_{ij}y_i y_j),}$

\raggedright
\setlength{\parindent}{2em}

\noindent where the second sum is over all $\frac{1}{2}n(n-1)$ pairs of $n$
species. In this model, $\alpha_{i}$ is an intercept term determining the amount
that the presence of species $i$ contributes to the log-probability of
$\vec{y}$; it directly controls the prevalence of species $i$. Similarly,
$\beta_{ij}$ is the amount that the co-occurrence of species $i$ and species $j$
contributes to the log-probability; it controls the conditional relationship
between two species, i.e. the probability that they will be found together,
after controlling for the other species in the network (Figure 2A, Figure 2B).
For example, if $\beta_{ij} = +2$, then each species' odds of occurrence would be
$e^2$ times higher when the other one is present (as compared with otherwise
equivalent sites). The relative probability of a presence-absence vector
increases when positively-associated species co-occur and decreases when
negatively-associated species co-occur. As a result, the model tends---all else
equal---to produce assemblages where many positively-associated species pairs
co-occur and few negatively-associated pairs do (just as an ecologist might
expect). When all else is *not* equal (e.g. Figure 1, where the presence of one
competitor is associated with release from another competitor), then predicting
species' overall co-occurrence rates would require normalizing the relative 
probabilities of possible $\vec{y}$ vectors so that they sum to one (Figure 2B).

\noindent \textbf{Estimating $\alpha$ and $\beta$ coefficients from presence-absence data.}
In the previous section, the values of $\alpha$ and $\beta$ were known and the
goal was to make predictions about possible species assemblages. In
practice, however, ecologists will often need to estimate the parameters from
an observed co-occurrence matrix (i.e. from a set of independent $\vec{y}$ vectors
indicating which species are present at each site on the landscape). When
Equation 1 can be normalized, one can find exact maximum likelihood estimates for
$\alpha$ and $\beta$ using numerical optimization. Fully-observed networks like the 
ones considered here have unimodal likelihood surfaces [@murphy_machine_2012], so 
these estimates will represent the global optimum. I used the rosalia package 
[@harris_rosalia_2015] for the R programming language [@r_core_team_r_2015] to 
calculate $p(\vec{y}; \alpha, \beta)$ and its gradients [@murphy_machine_2012]; 
the package passes these functions to the "BFGS" method in R's general-purpose 
optimizer, which then estimates the Markov network parameters.

\noindent \textbf{Simulated landscapes.}
I simulated several sets of landscapes using known parameters to evaluate
different statistical methods' performance. The first set of landscapes included
the three competing species shown in Figure 1. For each of 1000 replicates, I
generated a landscape with 100 sites by sampling from a probability distribution
defined by the figure's interaction coefficients (Appendix 1). Each of the
methods described below was then evaluated on its ability to correctly infer
that the two shrub species competed with one another, despite their frequent
co-occurrence.

I then simulated landscapes with up to 20 interacting species at 25, 200, or
1600 sites using three increasingly complex models (50 replicates for each
combination of size and model). The simplest set of simulated landscapes were
generated with Gibbs samples from Equation 1, and provided a test of methods'
ability to isolate direct species-level interactions from a web of indirect 
ones. For each replicate, I randomly drew the “true” $\beta$ coefficient
magnitudes an exponential distribution with rate 1 so that
most species pairs interacted negligibly but a few interactions were strong
enough to propagate through the newtwork. I randomly assigned 25% of the
interactions to be positive; the remainder were negative.

The next set of landscapes provided a way to assess each method's ability to 
identify direct interactions in the presence of environmental heterogeneity.
Here, the $\beta$ coefficients were calculated as above, but each species'
$\alpha$ value depended linearly on two normally-distributed environmental 
factors, which were drawn from independent Gaussians for each site. Once the 
local $\alpha$ values were calculated, independent Gibbs samplers determined 
the species composition for each site based on Equation 1.

In the third set of landscapes, I simulated each species' abundance (instead of
just their presence/absence), and interactions between species occurred on a 
per-capita basis. To avoid runaway mutualisms leading to infinite abundance,
all interaction coefficients were negative in these simulations. If a method
performs well on these landscapes, it would indicate some robustness to the
biolocical mechanisms underlying species' interactiosn with one another.

\noindent \textbf{Recovering species interactions from simulated data.}
I compared seven techniques for determining the sign and strength of the
associations between pairs of species from simulated data (Appendix 3).
First, I used the rosalia package [@harris_rosalia_2015] to fit Markov network
models, as described above. For the analyses with 20 species, a weak regularizer
(equivalent to a logistic prior with scale 2) ensured that the estimates were
always finite.

I also evaluated six alternative methods: five from the existing literature,
plus a novel combination of two of these methods. The first alternative 
interaction metric was the sample correlation between species' presence-absence
vectors, which summarizes their marginal association.  Next, I used partial 
correlations, which summarize species' conditional relationships. This approach, 
which is  closely related to linear regression, is common in molecular biology
[@friedman_sparse_2008], but is rare in ecology (see @albrecht_spatial_2001 and
@faisal_inferring_2010 for two exceptions). In the context of non-Gaussian data,
the partial correlation can be thought of as a computationally efficient
approximation to the full Markov network model [@loh_structure_2013]. Partial
correlations are undefined for landscapes with perfectly-correlated species
pairs, so I used the regularized estimate provided by the corpcor package’s 
`pcor.shrink` function with the default settings [@schafer_corpcor_2014].

The third alternative, generalized linear models (GLMs), also provide a
computationally efficient approximation to the Markov network
[@lee_learning_2012].  Following @faisal_inferring_2010,  I fit regularized
logistic regression models [@gelman_weakly_2008] for each species, using the
other species on the landscape as predictors. This produced two interaction
estimates for each species pair; one for the effect of species $i$ on species
$j$ and one for the reverse. These two coefficients are not identifiable from
the data, however [@schmidt_modeling_2012], so I used their average as an
overall measure of the pairwise relationship.

The next method used the Pairs software described in @gotelli_empirical_2009.
This program simulating new landscapes from a null model that retains the row 
and column sums of the original matrix [@strong_ecological_1984] and calculates
$Z$-scores to summarize a species pair's deviation from this null.

The last two estimators used the latent correlation matrix estimated by the
BayesComm package [@golding_bayescomm_2015] in order to evaluate the recent
claim that the correlation coefficients estimated by "joint species distribution
models" provide an accurate assessment of species’ pairwise interactions
[@pollock_understanding_2014; see also @harris_generating_2015]. In addition to
using the posterior mean correlation [@pollock_understanding_2014], I also used
the posterior mean *partial* correlation, which should control better for
indirect effects.

\noindent \textbf{Evaluating model performance.}
For the simulated landscapes based on Figure 1, I assessed whether each method's
test statistic indicated a positive or negative relationship between the two
shrubs (Appendix 1). For the null model (Pairs), I calculated statistical
significance using its $Z$-score. For the Markov network, I used the Hessian
matrix to generate approximate confidence intervals.

For the larger landscapes, I evaluated the relationship between each method's
estimates and the "true" interaction strengths. To ensure that the different
test statistics (e.g. correlations versus $Z$ scores) were on a common scale, I
rescaled them using linear regression through the origin. I then calculated the
proportion of variance explained for different combinations of model type and
landscape size (compared with a baseline model that assumed all interaction
strengths to be zero).

For the null model and the Markov network, the probability of rejecting the
null hypothesis of zero interaction was estimated across a range of "true"
interaction strengths using a generalized additive model with binomial responses
(reject versus not; Appendix 4). The Type I error rate was defined as this
model's estimated probability of rejecting the null hypothesis when the "true"
interaction was zero.

\noindent
\textbf{Results}

\noindent \textbf{Three species.}
As shown in Figure 1, the marginal relationship between the two shrub species
was positive---despite their competition for space at a mechanistic level---due
to indirect effects of the dominant tree species. As a result, the correlation
between these species was positive in 94% of replicates, and the
randomization-based null model falsely reported positive associations 100% of
the time. Worse, more than 98% of these false conclusions were statistically
significant. The partial correlation and Markov network estimates, on the other
hand, each correctly isolated the direct negative interaction between the shrubs
from their positive indirect interaction 94% of the time (although the
confidence intervals overlapped zero in most replicates).

\noindent
\textbf{Twenty species.}
In general, each model's performance was highest for large landscapes with
simple assembly rules and no environmental heterogeneity (Figure 3).
Despite some variability across contexts, the rank ordering across methods was
very consistent.  In particular, the four methods that controlled for indirect
effects (the Markov network, the generalized linear models, and the two partial
correlation-based methods) always matched or outperformed those that did not.
The Markov network consistently performed best of all. As anticipated by
@lee_learning_2012, generalized linear models closely
approximated the Markov network estimates (Figure 4A), especially when the data
sets were very large (Figure 3). As reviewed in @gotelli_empirical_2009,
however, most analyses in this field of ecology involve fewer than 50 sites;
in this context, the gap between the methods was larger. As shown in Appendix 4,
the standard errors associated with the estimates in Figure 3 are small (less
than 0.01), so the differences among methods should not be attributed to 
sampling error.

Of the methods that did not control for indirect effects, Figure 3 shows that
simple correlation coefficients provided a more reliable indicator of species'
true interaction strengths than either the joint species distribution model
(BayesComm) or the null model (Pairs). 95% of the variance the Pairs test 
statistic was explained by correlation coefficients (controlling for landscape
size; Figure 4B); much of the remaining variance is due to sampling error.

Finally, we can evaluate the models' statistical inferences (focusing on the
first two simulation types, where the interaction coefficients are easiest to
interpret).  The Markov network's Type I error rate depended on the 
simulation type: [[0.02]] for simulations that matched the model's assumptions,
versus [[0.14]] for simulations that included environmental heterogeneity (see 
Appendix 4 for confidence interval coverage across a range of $\beta_{ij}$ 
values). The null model's Type I error rates were [[0.30]] and [[0.50]] for the
constant and heterogeneous landscapes, respectively---far higher than the 
nominal 0.05 rate. Figure 4C shows, across a range of true interaction 
strengths, the probability that the null model or the Markov network will 
predict the wrong sign of the interaction with 95% confidence.  The null model 
makes such errors more than [[5]] times as often as the Markov network, even
though it only rejects the null hypothesis [[twice]] as often overall (Appendix
4). The Markov network's errors were also more concentrated around 0, so it 
never misclassified strong interactions like the null model did.

\noindent
\textbf{Discussion}

\noindent
The results presented above show that Markov networks can reliably recover
species' pairwise interactions from species composition data, even for cases where
environmental heterogeneity and indirect interactions cause ecologists' typical
null modeling approaches to reliably fail. Partial covariances and generalized
linear models can both provide computationally efficient approximations, but
with somewhat lower accuracy [especially for typically-sized data sets with small
numbers of sites; @gotelli_empirical_2009]. The difference in accuracy may be
larger for real data sets than for the simulated landscapes in Figure 3,
however; linear approximations to the Markov network make larger errors when
the interaction matrix is structured [e.g. due to guilds or trophic levels;
@loh_structure_2013]. Similarly, the separate generalized linear models
for each species can severely overfit in some cases [@lee_learning_2012]. The
full Markov network should thus be preferred to the approximations when
it is computationally tractable.

Compositional data only contains enough degrees of freedom to estimate one
interaction per species pair [@schmidt_modeling_2012], so none of these methods
can identify the exact nature of the pairwise interactions (e.g. which species
in a positively-associated pair is facilitating the other). To estimate
asymmetric interactions, such as commensalism or predation, ecologists could use
time series, behavioral observations, manipulative experiments, or natural
history. These other sources of information could also be used to augment the
likelihood function with a more informative prior distribution, reducing ecologists'
error and uncertainty relative to Figure 3's results.

Despite their limitations, Markov networks have enormous potential to improve
ecological understanding. In particular, they make many fewer errors than existing
approaches, and can make precise statements about the conditions where indirect
interactions will overwhelm direct ones. They also provide a simple answer to
the question of how competition should affect a species' overall prevalence,
which has important implications for community-level modeling
[@strong_ecological_1984]. Specifically, Equation 1 can be used to calculate the
expected prevalence of a species in the absence of biotic influences as
$e^\alpha/(e^{0} + e^\alpha)$. Competition's effect on prevalence can then be
estimated by comparing this value with the observed prevalence (e.g. comparing
Figure 2D with Figure 2C). This novel quantitative result conflicts with most of
our null models, which unreasonably assume that prevalence would be the exactly
same in the absence of competition as it is in the observed data
[@roughgarden_competition_1983].

Markov networks---particularly the Ising model for binary networks---are very
well understood, having been studied for nearly a century
[@cipra_introduction_1987]. Tapping into this framework would thus allow
ecologists to take advantage of into a vast set of existing discoveries and
techniques for dealing with indirect effects, stability, and alternative stable
states. Numerous extensions to the basic network are possible as well. For
example, the states of the interaction network can be modeled as a function of
the local abiotic environment [@lee_learning_2012], which would help incorporate
networks of biotic interactions into species distribution models
[@pollock_understanding_2014] and lead to a better understanding of the
interplay between biotic and abiotic effects on community structure.
Alternatively, models could allow one species to alter the relationship between
two other species [@tjelmeland_markov_1998; cf @bruno_inclusion_2003].

Finally, the results presented here have important implications for
ecologists' continued use of null models for studying species interactions. Null
and neutral models can be useful for clarifying our thinking
[@harris_occupancy_2011; @xiao_strong_2015], but deviations from a particular
null model must be interpreted with care [@roughgarden_competition_1983]. Even
in small networks with three species, it may simply not be possible to implicate
specific ecological processes like competition by rejecting a general-purpose
null [@gotelli_empirical_2009], especially when the test statistic is
effectively just a correlation coefficient (Figure 4B). When the non-null
backdrop is not controlled for, Type I error rates can skyrocket, the apparent
sign of the interaction can change, and null models can routinely produce
misleading inferences (Figure 1, Figure 4C, @gotelli_empirical_2009).

Controlling for indirect effects via simultaneous estimation of multiple
ecological parameters seems like a much more promising approach: to the extent
that the models' relative performance on real data sets is similar to the range
of results shown in Figure 3, scientists in this field could often triple
their explanatory power by switching from null models to Markov networks (or
increase it nearly as much with linear or generalized linear approximations).
Regardless of the methods ecologists ultimately choose, controlling for indirect
effects could clearly improve our understanding of species' direct effects on
one another and on community structure.

\noindent \textbf{Acknowledgements:}
This work benefited greatly from discussions with A. Sih, M. L. Baskett, R.
McElreath, R. J. Hijmans, A. C. Perry, and C. S. Tysor. Additionally, A. K.
Barner, E. Baldridge, E. P. White, D. Li, D. L. Miller, N. Golding, N. J.
Gotelli, C. F. Dormann, and two anonymous reviewers provided very helpful
feedback on the text. This research was partially supported by a Graduate
Research Fellowship from the US National Science Foundation and by the Gordon
and Betty Moore Foundation's Data-Driven Discovery Initiative through Grant
GBMF4563 to E. P. White.

\setlength{\parindent}{0cm}

\noindent \textbf{References:}

\setlength{\parindent}{-1em}
\setlength{\leftskip}{1em}
\setlength{\parskip}{0pt}

**Figure 1: A.** A small network of three competing species.
The tree (top) tends not to co-occur with either of the two shrub species, as
indicated by the strongly negative coefficient linking them. The two shrub
species also compete with one another, but more weakly (circled coefficient).
**B.** In spite of the competitive interactions between the two shrub species,
their shared tendency to occur in locations without trees makes their occurrence
vectors positively correlated (circled). **C.** Controlling for trees with a
conditional (all-else-equal) approach such as a partial covariance or a Markov
network leads to correct identification of the negative shrub-shrub interaction
(circled). See Appendix 1 and the results for "three species" for more details.

**Figure 2: A.** A small Markov network, defined by its $\alpha$ and
$\beta$ values.  The abiotic environment favors the occurrence of each species
($\alpha >0$), particularly species 2 ($\alpha_2 > \alpha_1$). The negative
$\beta_{12}$ coefficient is consistent with competition between the two species.
**B.** The coefficients determine the probabilities of all four possible
presence-absence combinations for Species 1 and Species 2. $\alpha_1$ is added
to the exponent whenever Species 1 is present ($y_1 = 1$), but not when it is
absent ($y_1 = 0$).  Similarly, the exponent includes $\alpha_2$ only when
species $2$ is present ($y_2 = 1$), and includes $\beta_{12}$ only when both are
present ($y_1y_2 = 1$). The normalizing constant $Z$, ensures that the four
probabilities sum to 1.  In this case, $Z$ is about 18.5.  **C.** The expected
frequencies of all possible co-occurrence patterns between the two species of
interest, as calculated in the previous panel. **D.** Without competition (i.e.
with $\beta_{12}=0$, each species would occur more often.

**Figure 3:** Proportion of variance in interaction coefficients explained by
each method versus number of sampled locations across the three simulation
types. For the null model (Pairs), two outliers with $|Z|>1000$ were manually
adjusted to $|Z|=50$ to mitigate their detrimental influence on $R^2$ (Appendix
5).

**Figure 4: A.** The Markov network's estimated interaction
coefficients were generally very similar to the GLM estimates. **B.** The null
model's estimates typically matched the (negative) correlation coefficient,
after controlling for landscape size. **C.** For any given interaction strength,
the null model was much more likely to misclassify its sign with 95% confidence
than the Markov network was.