### 2016-03-14 Harris_2015/show-results.Rmd

---
output: 
  pdf_document:
    keep_tex: true
    toc: true
title: "Appendix 4: Results"
subtitle: "Inferring species interactions from co-occurrence data with Markov networks"
author: David J. Harris
fontsize: 11pt
---

# A: Import packages and data

```{r, message = FALSE}
library(dplyr)
library(magrittr)
library(mgcv)
library(ggplot2)
library(tidyr)
library(knitr)
library(lme4)
```

## Import the results from Appendix 3:

```{r, cache = TRUE}
x = read.csv("estimates.csv", stringsAsFactors = FALSE)
x$simulation_type = gsub("[0-9]", "", x$rep_name)
```

## Import the results from the *Pairs* software:

```{r, cache = TRUE}

pairs_txt = readLines("fakedata/matrices/Pairs.txt")
library(stringr)

# Find areas of the data file that correspond
# to species pairs' results
beginnings = grep("Sp1", pairs_txt) + 1
ends = c(
  grep("^[^ ]", pairs_txt)[-1],
  length(pairs_txt) + 1
) - 1

partial_names = sapply(
  strsplit(grep("^>", pairs_txt, value = TRUE), " +"), 
  function(x) x[[3]]
)
filename_lines = grep("^>", pairs_txt)

# Sort a vector of alphanumeric strings by the numeric component
# as if they were integers.  For example, V20 should be larger 
# than V12, even though V12 comes first alphabetically
alnum_sort = function(x){
  raw = as.integer(gsub("[[:alpha:]]", "", x))
  x[order(raw)]
}

pairs_results = lapply(
  1:length(filename_lines),
  function(i){

    n_sites = as.integer(strsplit(partial_names[[i]], "-")[[1]][[1]])
    rep_name = strsplit(partial_names[[i]], "-")[[1]][[2]]

    # Find the line where the current data set is mentioned in
    # pairs.txt
    filename_line = filename_lines[i]

    # Which chunk of the data file corresponds to this file?
    chunk = min(which(beginnings > filename_line))

    # Split the chunk on whitespace.
    splitted = strsplit(pairs_txt[beginnings[chunk]:ends[chunk]], " +")

    # Pull out the corresponding chunk of the "x" data frame, based on n_sites 
    # and rep_name
    is_correct_sim = x$n_sites == n_sites & x$rep_name == rep_name
    x_subset = x[is_correct_sim & x$method == "correlation", ]

    # Pull out the species numbers and their Z-scores, then join to x_subset
    pairs_results = lapply(
      splitted,
      function(x){
        # in the x data frame, species 1 is always a lower number than species 2
        spp = alnum_sort(x[3:4])
        data.frame(
          sp1 = spp[1], 
          sp2 = spp[2], 
          z = x[14],
          stringsAsFactors = FALSE
        )
      }
    ) %>% 
      bind_rows %>%
      mutate(spp = paste(sp1, sp2, sep = "-"))

    n_spp = 20

    pairs_results$z = as.numeric(pairs_results$z)

    # Re-order the pairs_results to match the other methods
    m = matrix(NA, n_spp, n_spp)
    new_order = match(
      paste0("V", row(m)[upper.tri(m)], "-V", col(m)[upper.tri(m)]),
      pairs_results$spp
    )
    ordered_pairs_results = pairs_results[na.omit(new_order), ]

    ordered_pairs_results = ordered_pairs_results %>%
      filter(sp1 %in% c(x_subset$sp1) & sp2 %in% x_subset$sp2)

    x_subset$estimate = ordered_pairs_results$z
    x_subset$method = "null"

    x_subset
  }
) %>% bind_rows()

# Manually adjust the Z values less than -1000 so that these outliers
# won't completely dominate the analyses below
pairs_results$estimate[pairs_results$estimate < -1000] = -50

x = rbind(x, pairs_results)

```

# B: Compare model estimates

## Calculate model performance (R-squared):

For each combination of `method` and `simulation_type`, fit a linear model.
Then, for each combination of `method`, `simulation_type`, and `n_sites`,
report the proportion of variance explained by the corresponding linear model.

```{r, cache = FALSE}
resids = function(data){
  resid(lm(truth ~ estimate + 0, data = data))
}

result_summary = x %>% 
  group_by(method, simulation_type) %>%
  do(data.frame(., resids = resids(.))) %>%
  ungroup %>%
  group_by(method, simulation_type, n_sites) %>%
  summarise(r2 = 1 - sum(resids^2) / sum(truth^2))

# Set the ordering of the data frame based on R-squared
result_summary$method = reorder(result_summary$method, -result_summary$r2)
result_summary$simulation_type = reorder(
  result_summary$simulation_type, 
  -result_summary$r2
)

# Add a column that includes method name and its mean R-squared,
# for plotting purposes below
result_summary = result_summary %>% 
  group_by(method) %>% 
  summarise(mean_r2 = round(100 * mean(r2))) %>%
  mutate(method_r2 = paste0(method, " (0.", mean_r2, ")")) %>%
  select(method, method_r2) %>%
  inner_join(result_summary, "method")

# Rename the simulation types for clearer graph labels
result_summary$simulation_type_long = plyr::revalue(
  result_summary$simulation_type,
  c(no_env = "constant environment",
    env = "heterogeneous environment",
    abund = "abundance")
)

result_summary$method_r2 = reorder(result_summary$method_r2, -result_summary$r2)
```

#### Plot the regressions used above to calculate R^2

The plots below each correspond to one of the regressions performed above to
calculate R^2 for a combination of simulation method and estimation method.
The null model ("Pairs") was not included in these plots because its Z-scores
were frequently too large to fit on the same scale as the other methods' estimates.
Below, the thinner dashed lines represent the 1:1 line, where the expected estimate
matches the "true" value, and the thicker solid lines represent a linear regression
through the origin, as calculated above. In general, the four correlation- or
partial correlation-based methods had slopes that were steeper than the 1:1 line,
because their estimates were constrained between -1 and 1. The GLM and Markov network
usually had slopes that were closer to 1.

(Warnings of the form "Removed N rows containing missing values
(geom_smooth)" can be ignored: they are simply noting that the regression lines 
were cut off in some panels of the figure).

```{r, dev="png", dpi=500, cache = TRUE}
plot_regressions = function(type, title){
  x[sample.int(nrow(x)) , ] %>%
  subset(simulation_type == type & method != "null") %>% 
  ggplot(aes(x = estimate, y = truth)) +
    facet_wrap(~method) +
    ylim(min(x$truth), max(x$truth)) + 
    geom_point(aes(color = factor(n_sites)), size = 0.2) + 
    geom_smooth(method = "lm", color = "black", formula = y ~ 0 + x, 
                fullrange = TRUE) + 
    geom_abline(intercept = 0, slope = 1, linetype = 2) +
    theme_bw() + 
    ggtitle(title)
}

plot_regressions("no_env", "'Baseline' regressions")
plot_regressions("env", "'Environment' regressions")
plot_regressions("abund", "'Abundance' regressions")
```

#### Average the R-squared values across methods and simulation types

```{r}
result_summary %>% 
  group_by(method, simulation_type) %>% 
  summarise(mean(r2)) %>% 
  spread(simulation_type, `mean(r2)`) %>%
  kable(digits = 3)
```

## Save the finer-grained R-squared results as Figure 3

```{r}

legend_name = expression(Method~(mean~R^2))

pdf("manuscript-materials/figures/performance.pdf", width = 8, height = 2.5)
ggplot(result_summary, aes(x = n_sites, y = r2, col = method_r2, shape = method_r2)) + 
  facet_grid(~simulation_type_long) + 
  geom_line(size = .5) + 
  geom_point(size = 2.5, fill = "white") + 
  scale_shape_manual(values = c(16, 22, 17, 23, 18, 24, 15), name = legend_name) +
  geom_hline(yintercept = 0, size = 1/2) +
  geom_vline(xintercept = 0, size = 1) + 
  coord_cartesian(ylim = c(-.01, 0.76)) + 
  ylab(expression(R^2)) + 
  xlab("Number of sites (log scale)") + 
  scale_x_log10(breaks = unique(x$n_sites), limits = range(x$n_sites)) +
  theme_bw(base_size = 11) + 
  theme(panel.margin = grid::unit(1.25, "lines")) + 
  theme(panel.border = element_blank(), axis.line = element_blank()) + 
  theme(
    panel.grid.minor = element_blank(), 
    panel.grid.major.y = element_line(color = "lightgray", size = 1/4), 
    panel.grid.major.x = element_blank()
  ) + 
  theme(strip.background = element_blank(), legend.key = element_blank()) + 
  theme(plot.margin = grid::unit(c(.01, .01, .75, .1), "lines")) + 
  theme(
    axis.title.x = element_text(vjust = -0.2, size = 12), 
    axis.title.y = element_text(angle = 0, hjust = -.1, size = 12)
  ) + 
  scale_color_brewer(palette = "Dark2", name = legend_name)
dev.off()
```

## Estimate uncertainty in mean R-squared:

Fit a linear mixed model describing R-squared as a function of method, landscape
size, and simulation type.  Note the small standard errors associated with the
effect of estimation method.

```{r, cache = FALSE}
landscape_estimates = x %>% 
  group_by(method, simulation_type) %>%
  do(data.frame(., resids = resids(.))) %>%
  ungroup %>%
  group_by(method, simulation_type, rep_name, n_sites) %>%
  summarise(r2 = 1 - sum(resids^2) / sum(truth^2))

summary(
  lmer(
    r2 ~ method + n_sites + simulation_type + (1|rep_name), 
    data = landscape_estimates
  ),
  correlation = FALSE
)
```

# C: Inferential statistics

Note that, as mentioned in the main text, inferential statistics are only 
calculated for the `env` and `no_env` simulations (and not for the `abund`
simulations).

## Identify statistical significance for the Markov network and Pairs

For Pairs, compare the Z-scores with Gaussian quantiles for 95% coverage;
for the Markov network, use the approximate confidence intervals estimated in
Appendix 3.

```{r, cache = FALSE}

pairs_summary = x[x$method == "null" & x$simulation_type != "abund", ]
markov_summary = x[x$method == "Markov network" & x$simulation_type != "abund", ]

# Pairs's Z score is based on C-scores, which are positive when species are 
# disaggregated.  So significantly negative scores imply positive interactions
# and vice versa
pairs_summary$sig_pos = pairs_summary$estimate < qnorm(.025)
pairs_summary$sig_neg = pairs_summary$estimate > qnorm(.975)

# Markov network is significant when lower bound is above zero or lower bound is
# below zero
markov_summary$sig_pos = markov_summary$lower > 0
markov_summary$sig_neg = markov_summary$upper < 0
```

The `error_smoother` is a function that fits a kernel smoother model to 
estimate the probability that a model estimate will match some criterion (e.g. 
statistical significance) for some parameter as a function of that parameter's
"true" value.  This function is used in Figure 4C and in calculating Type I
error rates below.

```{r}

truth_seq = seq(min(markov_summary$truth), max(markov_summary$truth), length = 1000)

error_smoother = function(data, f, values = truth_seq){
  out = ksmooth(
    x = data$truth,
    y = as.numeric(f(data)),
    x.points = values,
    kernel = "normal",
    bandwidth = 0.25
  )

  out$y[is.na(out$y)] = 0

  out
}
```

## Create Figure 4

Panels A and B plot the point estimates or test statistics for different models 
against one another.

Panel C shows the probability of rejecting the null hypothesis (of no 
interaction between two species) in the opposite direction of the true value
(see main text).

```{r, cache = FALSE}

pdf("manuscript-materials/figures/error_rates.pdf", height = 8.5, width = 8.5/3)
par(mfrow = c(3, 1))

# Calculate p(confidently wrong) for each model as a function of
# the true values
confidently_wrong = function(data){
  (data$sig_pos & data$truth < 0) | (data$sig_neg & data$truth > 0)
}
smoothed_markov = error_smoother(markov_summary, confidently_wrong)
smoothed_pairs = error_smoother(pairs_summary, confidently_wrong)

# Compare estimates
spread_estimates = x %>%
  dplyr::select(-lower, -upper, -X) %>%
  spread(method, estimate) %>%
  na.omit()

# R-squared for null versus correlation
round(
  summary(lm(null ~ I(correlation*sqrt(n_sites)), data = spread_estimates))$r.squared, 
  2
)

# R-squared for glm markov network versus glm
round(summary(lm(`Markov network` ~ GLM, data = spread_estimates))$r.squared, 2)

with(
  spread_estimates, 
  plot(
    `Markov network`, 
    GLM, 
    pch = ".", 
    col = "#00000020",
    ylab = "Markov network estimate",
    xlab = "GLM estimate",
    bty = "l"
  )
)
mtext("A. Markov network estimates\nvs. GLM estimates", adj = 0, side = 3, 
      font = 2, line = 1.2, cex = .9) 
abline(lm(`Markov network` ~ GLM, data = spread_estimates))
text(0, 5, expression(R^2==0.94))

with(
  spread_estimates, 
  plot(
    correlation * sqrt(n_sites), 
    null, 
    pch = ".", 
    col = "#00000020",
    ylab = "Z-score",
    xlab = expression("correlation" %*% sqrt(number~~of~~sites)),
    bty = "l"
  )
)
mtext("B. Null model estimates vs.\nscaled correlation coefficients", 
      adj = 0, side = 3, font = 2, line = 1.2, cex = .9) 
abline(lm(null ~ I(correlation*sqrt(n_sites)), data = spread_estimates))
text(10, 12, expression(R^2==0.95))

plot(
  smoothed_pairs$x,
  smoothed_pairs$y,
  type = "l",
  xlab = "\"True\" interaction strength",
  ylab = "P(confidently predict wrong sign)",
  bty = "l",
  yaxs = "i",
  col = 2,
  ylim = c(0, .3),
  lwd = 2
)
mtext("C. Error rate vs.\ninteraction strength", side = 3, adj = 0, 
      font = 2, line = 1.2, cex = .9) 
lines(smoothed_markov$x, smoothed_markov$y, lwd = 2)
legend("topleft", lwd = 2, legend = c("Null model", "Markov network"), 
       col = c(2, 1), bty = "n")
dev.off()
```

## Summarize inferential statistics:

#### P(Pairs confidently wrong):
```{r}
with(pairs_summary,  mean((sig_neg & truth > 0) | (sig_pos & truth < 0)))
```

#### P(Markov network confidently wrong)
```{r}
with(markov_summary, mean((sig_neg & truth > 0) | (sig_pos & truth < 0)))
```

#### P(Pairs rejects null)
```{r}
with(pairs_summary,  mean(sig_neg | sig_pos))
```

#### P(Markov network rejects null)
```{r}
with(markov_summary,  mean(sig_neg | sig_pos))
```

## Type I error rates:

Type I error rates (probability of rejecting the null hypothesis given a 
"true" interaction strength of zero, according to the kernel smoother defined by
`error_smoother` above).

#### Markov network Type I error rates
```{r}
markov_summary %>% 
  filter(abs(truth) < 1) %>% 
  group_by(simulation_type) %>% 
  do(data.frame(
    `Type I error rate` = 
      error_smoother(., function(x){x$sig_pos | x$sig_neg}, values = 0)$y
  )) %>%
  kable(digits = 3)
```

#### Null model Type I error rates
```{r}
pairs_summary %>% 
  filter(abs(truth) < 1) %>% 
  group_by(simulation_type) %>% 
  do(data.frame(
    `Type I error rate` = 
      error_smoother(., function(x){x$sig_pos | x$sig_neg}, values = 0)$y
  )) %>%
  kable(digits = 3)
```

## Plot Markov network confidence interval coverage versus true $\beta$ value

The top and bottom 0.5% of the distribution have been omitted to prevent bad 
behavior by the smoother (a generalized additive model) in the tails. The 
horizontal red line shows the nominal 95% coverage rate and the black vertical 
line marks where the "true" value of $\beta$ was zero. The Type I error rates 
calculated above for the Markov network closely match (one minus) the values of 
these curves at zero.

```{r}
coverage_data = markov_summary %>%
  filter(percent_rank(truth) > .005 & percent_rank(truth) < .995) %>%
  mutate(covered = truth > lower & truth < upper) %>%
  mutate(simulation_type = factor(simulation_type, c("no_env", "env", "abund")))

ggplot(coverage_data, aes(x = truth, y = as.integer(covered))) + 
  facet_grid(~simulation_type) + 
  geom_smooth(method = gam, formula = y ~ s(x), method.args = list(family = binomial)) + 
  theme_bw() + 
  coord_cartesian(ylim = c(0, 1), expand = FALSE) + 
  geom_vline(xintercept = 0) + 
  geom_hline(yintercept = 0.95, color = "red") + 
  ylab("Coverage")

```

## Plot power versus true $\beta$ value

Calculate p(confidently detect sign of $\beta$) as a function of
the "true"" values.  Pairs has a somewhat higher rate of rejecting the null 
hypothesis in the correct direction than the Markov network, but the benefit of 
this increase in power is more than offset by the large increases in Type I 
errors and sign errors demonstrated above and in Figure 4 of the main text.

```{r}
confidently_right = function(data){
  (data$sig_pos & data$truth > 0) | (data$sig_neg & data$truth < 0)
}

plot(error_smoother(markov_summary, confidently_right), 
     type = "l", 
     xlim = c(-5, 5),
     xlab = "'True' coefficient value",
     ylab = "Power",
     ylim = c(0, 1),
     yaxs = "i"
)
lines(error_smoother(pairs_summary, confidently_right), col = 2)
legend("topleft", legend = c("Markov network", "Pairs"), lty = 1, col = 1:2)
```